---
title: "R Notebook"
author: "LMM"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r packages, warning = FALSE}
library(tidyverse) 
library(vegan)
library(compositions)
library(GGally)
library(ggrepel)
library(patchwork)
library(tidylog)
library(ALDEx2)
library(viridis)
```

```{r raw-data}
raw_asv <- read_csv("output/asv_raw.csv")

sample_list <- read_csv("data/sample_labels.csv") |> 
  filter(tissue != "dyctiota" & month == "october")

env <- read_csv("data/environmental.csv")
```

Some of the taxa can be resolved to a slightly better resolution and we filter chloroplasts out

```{r taxa-renaming}
raw_asv <- raw_asv |> 
  mutate(Family = ifelse(Class == "NB1-j", 
                        "NB1-j",
                        Family)) |> 
  mutate(Order = case_when(
    Class == "Marinimicrobia (SAR406 clade)" ~ "Deltaproteobacteria",
    Class == "NB1-j" ~ "Deltaproteobacteria",
    TRUE ~ Order)) |> 
  mutate(Class = case_when(
    Class == "Marinimicrobia (SAR406 clade)" ~ "Marinimicrobia",
    Class == "SAR324 clade(Marine group B)" ~ "Proteobacteria",
    Class == "NB1-j" ~ "Proteobacteria",
    Class == "WPS-2" ~ "Eremiobacterota (WPS-2)",
    is.na(Class) ~ "Unidentified Bacteria",
    TRUE ~ Class)) |> 
  filter(Family != "Chloroplast")
```


```{r count-data}
asv_taxa <- raw_asv |> 
  mutate(id = paste("asv", seq(1:nrow(raw_asv)), sep = "")) |>
  dplyr::select(c(1, 27:33))

asv_counts <- raw_asv |>
  dplyr::select(-c(1, 27:32)) |> #remove taxonomical information
  mutate(id = paste("asv", seq(1:nrow(raw_asv)), sep = "")) |> 
  pivot_longer(-id, names_to = "sample", values_to = "counts") |>  
  pivot_wider(values_from = counts, names_from = id) |> 
  mutate(sample = str_remove(sample, "_S.*"))
```

```{r taxa-names}
colnames(asv_taxa)

colnames(asv_taxa) <- c("ASV", "Domain", "Phylum", "Class", "Order", "Family", "Genus", "id")
```


Abundance of zeroes

```{r verification-zeroes}
raw_asv |>  
  #dplyr::select(where(is.numeric)) |>  
  rowwise() |> 
  mutate(total = sum(c_across(2:26))) |> 
 dplyr::select(total) |> 
  filter(total == 0)

rows_zeros <- asv_counts |>  
  #dplyr::select(where(is.numeric)) |>  
  rowwise() |> 
  mutate(total = sum(c_across(2:1897))) |>
  dplyr::select(sample, total) |> 
  filter(total == 0)

rows_zeros

asv_counts |>  
  #dplyr::select(where(is.numeric)) |>  
  rowwise() |> 
  mutate(total = sum(c_across(2:1897))) |>
  dplyr::select(total) |> 
  filter(total == 1)

checkNumZerosCol <- apply(asv_counts,2,function(x) sum(x==0))

cases <- which(checkNumZerosCol == (nrow(asv_counts) - 1))

length(cases) 
```

After removing chloroplasts, one row becomes all zero, which needs to be removed

```{r count-cleaning}
asv_counts <- asv_counts |> 
  filter(sample %in% sample_list$sample) |> 
  filter(!(sample %in% rows_zeros$sample))

asv_counts <- asv_counts[,-cases]
```

## Data transformation

One option would be to use the zCompositions::cmultRepl() method, however, in order to keep consistency, we can use the approach implemented in the Aldex2 package

```{r clr}
asv_props <- asv_counts |>  
  dplyr::select(where(is.numeric)) |> 
  mutate(across(everything(),
         ~.x +0.5)) |> 
  compositions::clr() #clr-normalization
```


```{r no-rhizomes}
lw_counts <- asv_counts |> 
  filter(!(str_detect(sample, "sgr")))

lw_props <- lw_counts |>  
  dplyr::select(where(is.numeric)) |> 
  mutate(across(everything(),
         ~.x +0.5)) |> 
  compositions::clr() #clr-normalization
```

### Permanova and permdisp

```{r permanova}
anova_factors <- data.frame(asv_counts$sample) |> 
  rename(sample = "asv_counts.sample") |> 
  left_join(sample_list) |> 
  tidylog::filter(!(sample %in% rows_zeros$sample)) |> 
  tidylog::filter(compartment != "rhizome") |> 
  tidylog::select(sample, treatment, compartment) 

anova_factors$interaction <-  paste(anova_factors$treatment, 
                                    anova_factors$compartment, sep = "-")

set.seed(1)
lw_props |> 
  data.frame() |>
  dplyr::select(where(is.numeric)) |> 
  (\(x) {adonis2(formula = x~anova_factors$treatment*anova_factors$compartment,
                method = "euclidean",
                permutations = 9999)})()
set.seed(1)
lw_props |> 
  data.frame() |>
  vegdist(method = "euclidean") |> 
  betadisper(group = anova_factors$interaction) |> 
  permutest()
```

# Taxa composition

```{r phyla-composition-bars-avg}
set.seed(3)
cols = turbo(26)[sample(1:26,26)]

lw_props |>
  compositions::clrInv() |>
  cbind(anova_factors[anova_factors$compartment!="rhizome",]) |>
  pivot_longer(names_to = "asv",
               values_to = "props",
               -c(compartment, treatment, sample, interaction)
               ) |>
  mutate(compartment = ifelse(compartment == "leaves", "Leaves", "Water column"),
         treatment = ifelse(treatment == "vent", "Vent", "Ambient")) |> 
  left_join(asv_taxa, by = c("asv" = "id")) |>
  dplyr::select(props, sample, Phylum, treatment, compartment) |>
  group_by(sample, Phylum, treatment, compartment) |>
  summarise(total = sum(props), weight = n()) |>
  group_by(Phylum, treatment, compartment) |>
  summarise(avg = mean(total)) |>
  ggplot() +
  geom_bar(aes(y = avg, x = treatment, fill = Phylum),
           position = "stack", stat = "identity") +
  facet_grid(~compartment) +
  labs(y = "Proportion") +
  scale_fill_manual(values = cols) +
  theme_minimal() +
  #scale_fill_viridis(discrete = TRUE, option = "turbo") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") 
```

```{r class-composition-bars-avg}
set.seed(3)
cols = turbo(39)[sample(1:39,39)]

lw_props |>
  compositions::clrInv() |>
  cbind(anova_factors[anova_factors$compartment!="rhizome",]) |>
  pivot_longer(names_to = "asv",
               values_to = "props",
               -c(compartment, treatment, sample, interaction)
               ) |>
  mutate(compartment = ifelse(compartment == "leaves", "Leaves", "Water column"),
         treatment = ifelse(treatment == "vent", "Vent", "Ambient")) |> 
  left_join(asv_taxa, by = c("asv" = "id")) |>
  dplyr::select(props, sample, Class, treatment, compartment) |>
  group_by(sample, Class, treatment, compartment) |>
  summarise(total = sum(props), weight = n()) |>
  group_by(Class, treatment, compartment) |>
  summarise(avg = mean(total)) |>
  ggplot() +
  geom_bar(aes(y = avg, x = treatment, fill = Class),
           position = "stack", stat = "identity") +
  facet_grid(~compartment) +
  labs(y = "Proportion", x = "pH regime") +
  scale_fill_manual(values = cols) +
  theme_minimal() +
  #scale_fill_viridis(discrete = TRUE, option = "turbo") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") 
```



### Ordination

```{r no-rhizomes}
lw_counts <- asv_counts |> 
  filter(!(str_detect(sample, "sgr")))

lw_props <- lw_counts |>  
  dplyr::select(where(is.numeric)) |> 
  mutate(across(everything(),
         ~.x +0.5)) |> 
  compositions::clr() #clr-normalization
```

```{r pco-no-rhizome}
pco_lw <- lw_props |> 
  vegdist(method = "euclidean") |> 
  labdsv::pco(k = 3)
  
x_expv_lw <- round(100*(pco_lw$eig[c(1)]/sum(pco_lw$eig)))
y_expv_lw <- round(100*(pco_lw$eig[c(2)]/sum(pco_lw$eig)))

pco_plot_lw <- pco_lw$points |> 
  data.frame() |> 
  mutate(Compartment = anova_factors[anova_factors$compartment!="rhizome",]$compartment,
         Treatment = anova_factors[anova_factors$compartment!="rhizome",]$treatment) |>
  ggplot() +
  geom_point(aes(x = X1, y = X2, 
                 shape = Compartment, 
                 color = Treatment
                 ),
             size = 5) +
  labs(x = paste("PCo 1 (",
                 x_expv_lw,
                 "% of explained variance)",
                 sep = ""),
       y = paste("PCo 2 (",
                 y_expv_lw,
                 "% of explained variance)",
                 sep = ""),
       shape = "Compartment",
       color = "pH") + 
  scale_shape_discrete(labels = c("Leaf", "Water column")) +
  scale_color_discrete(labels = c("Vent", "Ambient")) +
  theme_bw() +
  theme(legend.position = "bottom")

pco_plot_lw

#ggsave("figures/fig_06_pco.svg")

round(100*(pco_lw$eig[c(1,2)]/sum(pco_lw$eig)))
```

# Differential abundance analysis
## Seagrass vs water column

```{r DA-compartment}
covariates <- anova_factors |>
  dplyr::select(sample , compartment) |>
  arrange(sample) |>
  mutate(value = 1) |>
  pivot_wider(names_from = compartment, values_from = value, values_fill = 0) |>
  dplyr::pull(leaves)

aldex_output <- lw_counts |> 
  pivot_longer(names_to = "asv", values_to = "counts", -sample) |> 
  dplyr::filter(!(str_detect(sample, "sgr"))) |> 
  pivot_wider(names_from = sample, values_from = counts) |> 
  column_to_rownames("asv") |>
  aldex.clr(conds = covariates,
            verbose = TRUE,
            mc.samples = 1000,
            denom = "median")

aldex_tt <- aldex.ttest(aldex_output, verbose = TRUE)
aldex_eff <- aldex.effect(aldex_output, CI = TRUE)

aldex_full <- data.frame(aldex_tt, aldex_eff)
```


```{r}
aldex_full |> 
  mutate(significance = case_when(
    we.eBH < 0.05 & aldex_full$wi.eBH < 0.05 ~ "all",
    we.eBH < 0.05 | aldex_full$wi.eBH < 0.05 ~ "one",
    TRUE ~ "none"),
        slope = diff.win/diff.btw
    ) |> 
  ggplot() +
  geom_point(aes(x = diff.win, y = diff.btw, color = significance)) +
  geom_abline(slope = 1) +
  geom_abline(slope = -1) +
  labs(x = "Dispersion", y = "Difference", title="Effect size plot") +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "bottom")

#ggsave("supplementary/supp_1_aldex.svg")
```

```{r}
set.seed(3)
cols = turbo(17)[sample(1:17,17)]

aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Family, Genus, diff.btw, diff.win)) |> 
  mutate(Leaf = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column"))|> #View()
  group_by(Phylum, Class, Leaf) |> 
  summarise(mean_diff=mean(diff.btw), 
            mean_disp=mean(diff.win), 
            Total=n())|>
  ggplot() +
  geom_point(aes(y = reorder(Class, mean_diff), 
                 x = Leaf, size = mean_diff, 
                 color = Phylum)) +
  labs(x = "",
       y = "Class") +
  scale_color_manual(values = cols) +
  scale_size_area() +
  theme_bw() +
  guides(color = guide_legend(override.aes = list(size=5)))

#ggsave("figures/fig_7.svg")
```