---
title: "Accelerated nitrogen cycling on seagrass leaves in a high-CO2 world"
author: "LMM"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r packages, warning = FALSE}
library(tidyverse) 
library(vegan)
library(compositions)
library(GGally)
library(ggrepel)
library(patchwork)
library(tidylog)
library(ALDEx2)
library(viridis)
library(knitr)
library(ggh4x)
```

```{r raw-data}
raw_asv <- read_csv("output/asv_raw.csv")

sample_list <- read_csv("data/sample_labels.csv") 

env <- read_csv("data/environmental.csv")
```

Some of the taxa can be resolved to a slightly better resolution and we filter chloroplasts out

```{r taxa-renaming}
raw_asv <- raw_asv |> 
  mutate(Family = ifelse(Class == "NB1-j", 
                        "NB1-j",
                        Family)) |> 
  mutate(Order = case_when(
    Class == "Marinimicrobia (SAR406 clade)" ~ "Deltaproteobacteria",
    Class == "NB1-j" ~ "Deltaproteobacteria",
    TRUE ~ Order)) |> 
  mutate(Class = case_when(
    Class == "Marinimicrobia (SAR406 clade)" ~ "Marinimicrobia",
    Class == "SAR324 clade(Marine group B)" ~ "Proteobacteria",
    Class == "NB1-j" ~ "Proteobacteria",
    Class == "WPS-2" ~ "Eremiobacterota (WPS-2)",
    is.na(Class) ~ "Unidentified Bacteria",
    TRUE ~ Class)) |> 
  filter(Family != "Chloroplast")
```

Having this, we can generate the count data:

```{r count-data}
asv_taxa <- raw_asv |> 
  mutate(id = paste("asv", seq(1:nrow(raw_asv)), sep = "")) |>
  dplyr::select(c(1, 19:25))

asv_counts <- raw_asv |>
  dplyr::select(-c(1, 19:24)) |> #remove taxonomical information
  mutate(id = paste("asv", seq(1:nrow(raw_asv)), sep = "")) |> 
  pivot_longer(-id, names_to = "sample", values_to = "counts") |>  
  pivot_wider(values_from = counts, names_from = id) |> 
  mutate(sample = str_remove(sample, "_S.*"))
```

```{r taxa-names}
colnames(asv_taxa)

colnames(asv_taxa) <- c("ASV", "Domain", "Phylum", "Class", "Order", "Family", "Genus", "id")
```

Abundance of zeroes

```{r verification-zeroes}
raw_asv |>  
  #dplyr::select(where(is.numeric)) |>  
  rowwise() |> 
  mutate(total = sum(c_across(2:18))) |>
  dplyr::select(total) |> 
  filter(total == 0)

rows_zeros <- asv_counts |>  
  #dplyr::select(where(is.numeric)) |>  
  rowwise() |> 
  mutate(total = sum(c_across(2:1897))) |>
  dplyr::select(sample, total) |> 
  filter(total == 0)

rows_zeros

asv_counts |>  
  #dplyr::select(where(is.numeric)) |>  
  rowwise() |> 
  mutate(total = sum(c_across(2:1897))) |>
  dplyr::select(total) |> 
  filter(total == 1)

checkNumZerosCol <- apply(asv_counts,2,function(x) sum(x==0))

cases <- which(checkNumZerosCol == (nrow(asv_counts) - 1))

length(cases) 
```

After removing chloroplasts, one row becomes all zero, which needs to be removed

```{r count-cleaning}
asv_counts <- asv_counts |> 
  filter(sample %in% sample_list$sample) |> 
  filter(!(sample %in% rows_zeros$sample))

asv_counts <- asv_counts[,-cases]
```

## Data transformation

One option would be to use the zCompositions::cmultRepl() method, however, in order to keep consistency, we can use the approach implemented in the Aldex2 package

```{r clr}
lw_counts <- asv_counts |> 
  filter(!(str_detect(sample, "sgr")))

lw_props <- lw_counts |>  
  dplyr::select(where(is.numeric)) |> 
  mutate(across(everything(),
         ~.x +0.5)) |> 
  compositions::clr() #clr-normalization
```

### Permanova and permdisp

```{r permanova}
anova_factors <- data.frame(asv_counts$sample) |> 
  rename(sample = "asv_counts.sample") |> 
  left_join(sample_list) |> 
  tidylog::filter(!(sample %in% rows_zeros$sample)) |> 
  tidylog::filter(compartment != "rhizome") |> 
  tidylog::select(sample, treatment, compartment) 

anova_factors$interaction <-  paste(anova_factors$treatment, 
                                    anova_factors$compartment, sep = "-")

set.seed(1)
aov1 <- lw_props |> 
  data.frame() |>
  dplyr::select(where(is.numeric)) |> 
  (\(x) {adonis2(formula = x~anova_factors$treatment*anova_factors$compartment,
                method = "euclidean",
                permutations = 9999)})()
set.seed(1)
lw_props |> 
  data.frame() |>
  vegdist(method = "euclidean") |> 
  betadisper(group = anova_factors$interaction) |> 
  permutest()
```

```{r supp-table-1}
sink("supplementary/supp_table_1.md")
aov1 |> 
  rownames_to_column("SoV") |> 
  mutate(SoV = c("pH regime",
                 "Compartment",
                 "Treatment x Compartment",
                 "Residual",
                 "Total")) |> 
  kable(format = "markdown",
        col.names = c("Source of variation",
                      "Degress of freedom",
                      "Sum of squares",
                      "R2",
                      "Pseudo-F",
                      "P(>F)"),
        caption = "Permutation-based analysis of variance of the microbial communities associated with P. oceanica leaves, water column and pH regime.")
sink()
```

# Taxa composition

```{r fig1-dominant-class-comp-avg}
cols = c(turbo(12), "#5d5d5d")
#cols = ggmnmlab::martink_pals$martink15

test <- lw_props |>
  compositions::clrInv() |>
  cbind(anova_factors[anova_factors$compartment!="rhizome",]) |>
  pivot_longer(names_to = "asv",
               values_to = "props",
               -c(compartment, treatment, sample, interaction)
               ) |>
  mutate(compartment = ifelse(compartment == "leaves", "Leaves", "Water column"),
         treatment = ifelse(treatment == "vent", "Vent", "Ambient")) |> 
  left_join(asv_taxa, by = c("asv" = "id")) |>
  dplyr::select(props, sample, Phylum, Class, Order, Family, treatment, compartment) |>
  group_by(sample, Phylum, Class, Order, Family, treatment, compartment) |>
  summarise(total = sum(props), weight = n()) |>
  group_by(Phylum, Class, Order, Family, treatment, compartment) |>
  summarise(avg = mean(total)) |>
  mutate(threshold = ifelse(avg > 0.01, 1, 0)) 


f1a <- test |> 
  mutate(avg = avg*100,
         Class = ifelse(threshold == 0, "Other (RA<1%)", Class),
         Phylum = ifelse(threshold == 0, "Other (RA<1%)", Phylum)) |> 
  filter(threshold == 1 | Class == "Other (RA<1%)") |> 
  group_by(Phylum, Class, treatment, compartment) |>
  summarise(avg = sum(avg)) |>
  ggplot() +
  geom_bar(aes(y = avg, 
               x = treatment, 
               fill = fct_relevel(Class, "Other (RA<1%)", after = Inf)),
           position = "stack", 
           stat = "identity", 
           color = "black") +
  facet_grid(~compartment) +
  labs(y = "Relative abundance (%)", 
       x = "pH regime",
       tag = "B") +
  #scale_fill_viridis(option = "H", discrete = TRUE, name = "Class") +
  #scale_fill_mnmlab(palette = "martink15", name = "Class") +
  scale_fill_manual(values = cols, name = "Class") +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 15),
        #axis.title.y = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 11),
        plot.tag = element_text(size = 20))

f1b <- test |> 
  mutate(avg = avg*100,
         Phylum = ifelse(threshold == 0, "Other (RA<1%)", Phylum)) |> 
  filter(threshold == 1 | Phylum == "Other (RA<1%)") |> 
  group_by(Phylum, treatment, compartment) |>
  summarise(avg = sum(avg)) |>
  ggplot() +
  geom_bar(aes(y = avg, 
               x = treatment, 
               fill = fct_relevel(Phylum, "Other (RA<1%)", after = Inf)),
           position = "stack", 
           stat = "identity", 
           color = "black") +
  facet_grid(~compartment) +
  labs(x = "pH regime", tag = "A") +
  scale_fill_viridis(option = "H", 
                     discrete = TRUE, 
                     name = "Phylum") +
  #scale_fill_mnmlab(palette = "martink15", name = "Class") +
  #scale_fill_manual(values = cols, name = "Order") +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 15),
        #axis.title.y = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 11),
        plot.tag = element_text(size = 20))

f1c <- test |> 
  mutate(avg = avg*100,
         Family = ifelse(threshold == 0, "Other (RA<1%)", Family),
         Order = ifelse(threshold == 0, "Other (RA<1%)", Order),
         Class = ifelse(threshold == 0, "Other (RA<1%)", Class),
         Phylum = ifelse(threshold == 0, "Other (RA<1%)", Phylum)) |> 
  filter(threshold == 1 | Order == "Other (RA<1%)") |> 
  group_by(Phylum, Class, Order, Family, treatment, compartment) |>
  summarise(avg = sum(avg)) |> 
  mutate(Family = ifelse(is.na(Family), 
                         paste("Unidentified", Order), Family)) |> 
 ggplot() +
  geom_bar(aes(y = avg, x = treatment, 
               fill = fct_relevel(Family, "Other (RA<1%)", after = Inf)),
           position = "stack", stat = "identity", color = "black") +
  facet_grid(~compartment) +
  labs(x = "pH regime", tag = "C") +
  scale_fill_viridis(option = "H", 
                     discrete = TRUE, 
                     name = "Family") +
  #scale_fill_mnmlab(palette = "martink15", name = "Class") +
  #scale_fill_manual(values = cols, name = "Order") +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 15),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 11),
        plot.tag = element_text(size = 20))

f1b + f1a + f1c
ggsave("figures/fig1.svg", width = 24)
```

```{r fig2}
f2 <- lw_props |>
  compositions::clrInv() |>
  cbind(anova_factors[anova_factors$compartment!="rhizome",]) |>
  pivot_longer(names_to = "asv",
               values_to = "props",
               -c(compartment, treatment, sample, interaction)
               ) |>
  mutate(compartment = ifelse(compartment == "leaves", "Leaves", "Water column"),
         treatment = ifelse(treatment == "vent", "Vent", "Ambient")) |> 
  left_join(asv_taxa, by = c("asv" = "id")) |>
  filter(str_starts(Family, "Nitro")) |>
  group_by(sample, Family, treatment, compartment) |>
  summarise(total = sum(props), weight = n()) |>
  group_by(Family, treatment, compartment) |>
  summarise(avg = mean(total)) |>
  mutate(avg = avg*100) |> 
  ggplot() +
  geom_bar(aes(y = avg, x = treatment, fill = Family),
           position = "stack", stat = "identity",color = "black") +
  facet_grid(~compartment) +
  labs(y = "Relative abundance (%)", x = "pH regime") +
  scale_fill_viridis(option = "H", discrete = TRUE) +
  theme_minimal() +
  #scale_fill_viridis(discrete = TRUE, option = "turbo") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") 

#ggsave(fig2, "figures/figure2.svg")
```

We can separately explore the rare taxa as stacked bars

```{r}
test |> 
  filter(threshold == 0) |> 
  mutate(avg = avg*100) |> 
 ggplot() +
  geom_bar(aes(y = avg, x = treatment, fill = Class),
           position = "stack", stat = "identity", color = "black") +
  facet_grid(~compartment) +
  labs(y = "Relative abundance (%)", x = "pH regime", tag = "a") +
  scale_fill_viridis(option = "H", discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom",
        aspect.ratio = 2/1,
        strip.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 11),
        plot.tag = element_text(size = 20))
```

Or as a bubble matrix

```{r}
test |> 
  filter(threshold == 0) |> 
 ggplot() +
  geom_point(aes(size = avg, x = treatment, y = Class, color = Phylum)) +
  facet_grid(~compartment) +
  labs(y = "Proportion", x = "pH regime", tag = "a") +
  scale_fill_viridis(option = "H", discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom",
        aspect.ratio = 2/1,
        strip.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 11),
        plot.tag = element_text(size = 20))
```

...or as a table

```{r}
test |> 
  filter(threshold == 0) |> 
  mutate(avg = avg*100) |> 
 ggplot() +
  geom_tile(aes(fill = avg, x = treatment, y = Class)) +
  geom_text(aes(label = (round(avg, 2)), x = treatment, y = Class)) +
  facet_grid(~compartment) +
  labs(y = "Taxonomic class", x = "pH regime", tag = "b") +
  scale_fill_gradient(low = "white", high = "green")  +
  theme_minimal() +
  guides(fill = "none") +
  scale_y_discrete(limits = rev) +
  theme(legend.position = "bottom",
        aspect.ratio = 2/1,
        strip.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 11),
        plot.tag = element_text(size = 20))
```


```{r}
f1c
#ggsave("figures/fig2.svg", width = 7, height = 7)
```


```{r aov-2}
anova_factors_ntr <- lw_props |>
  compositions::clrInv() |>
  cbind(anova_factors[anova_factors$compartment!="rhizome",]) |>
  pivot_longer(names_to = "asv",
               values_to = "props",
               -c(compartment, treatment, sample, interaction)
               ) |>
  mutate(compartment = ifelse(compartment == "leaves", "Leaves", "Water column"),
         treatment = ifelse(treatment == "vent", "Vent", "Ambient")) |> 
  left_join(asv_taxa, by = c("asv" = "id")) |>
  filter(str_starts(Family, "Nitro")) |> 
  dplyr::select(compartment, treatment, interaction) 

aov2 <- lw_props |>
  compositions::clrInv() |>
  cbind(anova_factors[anova_factors$compartment!="rhizome",]) |>
  pivot_longer(names_to = "asv",
               values_to = "props",
               -c(compartment, treatment, sample, interaction)
               ) |>
  mutate(compartment = ifelse(compartment == "leaves", "Leaves", "Water column"),
         treatment = ifelse(treatment == "vent", "Vent", "Ambient")) |> 
  left_join(asv_taxa, by = c("asv" = "id")) |>
  filter(str_starts(Family, "Nitro")) |> 
  dplyr::select(-c(interaction, ASV, Domain, Phylum, Class, Order, Genus)) |> 
  pivot_wider(
              values_from = props,
              names_from = Family,
              values_fill = 0) |> 
  dplyr::select(where(is.numeric)) |> 
  (\(x) {adonis2(formula = x~anova_factors_ntr$treatment*anova_factors_ntr$compartment,
                method = "euclidean",
                permutations = 9999)})()
```


```{r supp-table-1}
sink("supplementary/supp_table_2.md")
aov2 |> 
  rownames_to_column("SoV") |> 
  mutate(SoV = c("pH regime",
                 "Compartment",
                 "Treatment x Compartment",
                 "Residual",
                 "Total")) |> 
  kable(format = "markdown",
        col.names = c("Source of variation",
                      "Degress of freedom",
                      "Sum of squares",
                      "R2",
                      "Pseudo-F",
                      "P(>F)"),
        caption = "Permutation-based analysis of variance of the nitrifying communities associated with P. oceanica leaves, water column and pH regime.")
sink()
```

### Ordination

```{r no-rhizomes}
lw_counts <- asv_counts |> 
  filter(!(str_detect(sample, "sgr")))

lw_props <- lw_counts |>  
  dplyr::select(where(is.numeric)) |> 
  mutate(across(everything(),
         ~.x +0.5)) |> 
  compositions::clr() #clr-normalization
```

```{r fig7-pco}
pco_lw <- lw_props |> 
  vegdist(method = "euclidean") |> 
  labdsv::pco(k = 3)
  
x_expv_lw <- round(100*(pco_lw$eig[c(1)]/sum(pco_lw$eig)))
y_expv_lw <- round(100*(pco_lw$eig[c(2)]/sum(pco_lw$eig)))

pco_plot_lw <- pco_lw$points |> 
  data.frame() |> 
  mutate(Compartment = anova_factors[anova_factors$compartment!="rhizome",]$compartment,
         Treatment = anova_factors[anova_factors$compartment!="rhizome",]$treatment) |>
  ggplot() +
  geom_point(aes(x = X1, y = X2, 
                 shape = Compartment, 
                 color = Treatment
                 ),
             size = 5) +
  labs(x = paste("PCo 1 (",
                 x_expv_lw,
                 "% of explained variance)",
                 sep = ""),
       y = paste("PCo 2 (",
                 y_expv_lw,
                 "% of explained variance)",
                 sep = ""),
       shape = "Compartment",
       color = "pH") + 
  scale_shape_discrete(labels = c("Leaf", "Water column")) +
  scale_color_discrete(labels = c("Vent", "Ambient")) +
  theme_bw() +
  theme(legend.position = "bottom")

pco_plot_lw

#ggsave("supplementary/supp_fig_01.svg")

round(100*(pco_lw$eig[c(1,2)]/sum(pco_lw$eig)))
```

# Differential abundance analysis
## Seagrass vs water column

First let's run the algorithm

```{r DA-compartment}
covariates <- anova_factors |>
  dplyr::select(sample , compartment) |>
  arrange(sample) |>
  mutate(value = 1) |>
  pivot_wider(names_from = compartment, values_from = value, values_fill = 0) |>
  dplyr::pull(leaves)

aldex_output <- lw_counts |> 
  pivot_longer(names_to = "asv", values_to = "counts", -sample) |> 
  dplyr::filter(!(str_detect(sample, "sgr"))) |> 
  pivot_wider(names_from = sample, values_from = counts) |> 
  column_to_rownames("asv") |>
  aldex.clr(conds = covariates,
            verbose = TRUE,
            mc.samples = 1000,
            denom = "median")

aldex_tt <- aldex.ttest(aldex_output, verbose = TRUE)
aldex_eff <- aldex.effect(aldex_output, CI = TRUE)

aldex_full <- data.frame(aldex_tt, aldex_eff)
```


```{r}
aldex_full |> 
  mutate(significance = case_when(
    we.eBH < 0.05 & aldex_full$wi.eBH < 0.05 ~ "all",
    we.eBH < 0.05 | aldex_full$wi.eBH < 0.05 ~ "one",
    TRUE ~ "none"),
        slope = diff.win/diff.btw
    ) |> 
  ggplot() +
  geom_point(aes(x = diff.win, y = diff.btw, color = significance)) +
  geom_abline(slope = 1) +
  geom_abline(slope = -1) +
  labs(x = "Dispersion", y = "Difference", title="Effect size plot") +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "bottom")

#ggsave("supplementary/supp_x_aldex.svg")
```

### Visualization

First approach: as bubble plots of the effect size

```{r fig8-diff-abund}
set.seed(3)
 cols = c(viridis(6), viridis(6), viridis(5))
 shp <- c(rep("\u25CF", 6), rep("\u25C9", 6), rep("\u25CD", 5))
 
aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Order, Family, Genus, diff.btw, diff.win)) |> 
  mutate(Compartment = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column"))|>
  group_by(Phylum, Class, Compartment) |> 
  summarise(mean_diff=mean(diff.btw), 
            mean_disp=mean(diff.win), 
            Total=n())|>
  ggplot() +
  geom_point(aes(y = reorder(Class, mean_diff), 
                 x = Compartment, 
                 size = mean_diff, 
                 color = Phylum,
                 shape = Phylum
                 )) +
  labs(x = "",
       y = "Class") +
  scale_color_manual(values = cols) +
  #scale_color_viridis(option = "H", discrete = TRUE) +
  scale_size_area(max_size = 9) +
  scale_shape_manual(values = shp) +
  theme_bw() +
  guides(color = guide_legend(override.aes = list(size=5)),
         size = "none")

#ggsave("figures/fig3.svg", width = 5, height = 7)
```

As barplots, by Family

```{r}
aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Order, Genus, diff.btw, diff.win)) |> 
  mutate(Compartment = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column"))|>
  group_by(Phylum, Class, Order, Compartment) |> 
  summarise(mean_diff=mean(diff.btw), 
            mean_disp=mean(diff.win), 
            Total=n()) |> 
  ggplot() +
  geom_col(aes(x = mean_diff, 
                 y = Order ,
                 fill = Compartment
                 )) +
  geom_vline(xintercept = 0) +
  scale_y_discrete(limits=rev) +
  labs(x = "Differential abundance") +
  theme_light() +
  theme(legend.position = "bottom") 
```


```{r}
aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Order, diff.btw, diff.win)) |> 
  mutate(Compartment = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column"))|>
  group_by(Phylum, Class, Compartment) |> 
  summarise(mean_diff=mean(diff.btw), 
            mean_disp=mean(diff.win), 
            Total=n()) |> 
  ggplot() +
  geom_col(aes(x = mean_diff, 
               y = Class ,
               fill = Compartment
  )) +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = c("#00BA38","#619CFF")) +
  facet_nested(
    rows = vars(Phylum),
    scales = "free",
    space = 'free',
    switch = "y",
    nest_line = element_line(color = "black",
                             linetype = 2)
  ) +
  labs(x = "Differential abundance",
       y = "Phylum & Class") +
  theme_light() +
  theme(legend.position = "bottom",
        strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0,
                                         colour = "black"),
        ggh4x.facet.nestline = element_line(colour = "blue",),
        strip.background = element_blank()
  )
```

```{r}
aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Order, diff.btw, diff.win)) |> 
  mutate(Compartment = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column"))|>
  group_by(Phylum, Class, Order, Compartment) |> 
  summarise(mean_diff=mean(diff.btw), 
            mean_disp=mean(diff.win), 
            Total=n()) |> 
  ggplot() +
  geom_col(aes(x = mean_diff, 
               y = Order ,
               fill = Compartment
  )) +
   scale_fill_manual(values = c("#00BA38","#619CFF")) +
  facet_nested(
    rows = vars(Phylum, Class),
    scales = "free",
    space = 'free',
    switch = "y",
    nest_line = element_line(color = "black",
                             linetype = 2)
  ) +
  labs(x = "Differential abundance",
       y = "Class & Order") +
  theme_light() +
  theme(legend.position = "bottom",
        strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0,
                                         colour = "black"),
        ggh4x.facet.nestline = element_line(colour = "blue",),
        strip.background = element_blank()
  )
```

```{r}
aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Order, diff.btw, diff.win)) |> 
  mutate(Compartment = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column"))|>
  group_by(Phylum, Class, Order, Compartment) |> 
  summarise(mean_diff=mean(diff.btw), 
            mean_disp=mean(diff.win), 
            Total=n()) |> 
  filter(Class %in% c("Alphaproteobacteria",
                      "Gammaproteobacteria",
                      "Cyanobacteriia",
                      "Rhodothermia",
                      "Bacteroidia",
                      "Verrucomicrobiae")) |> 
  ggplot() +
  geom_col(aes(x = mean_diff, 
               y = Order ,
               fill = Compartment
  )) +
  geom_vline(xintercept = 0) +
   scale_fill_manual(values = c("#00BA38","#619CFF")) +
  facet_nested(
    rows = vars(Phylum, Class),
    scales = "free",
    space = 'free',
    switch = "y",
    nest_line = element_line(color = "black",
                             linetype = 2)
  ) +
  labs(x = "Differential abundance",
       y =  "Class & Order") +
  theme_light() +
  theme(legend.position = "bottom",
        strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0,
                                         colour = "black"),
        ggh4x.facet.nestline = element_line(colour = "blue",),
        strip.background = element_blank()
  )
```

the highest resolution

```{r}
aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Order, Family, Genus, diff.btw, diff.win)) |> 
  mutate(Compartment = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column"),
         Family = ifelse(is.na(Family), Order, Family),
         Genus = ifelse(is.na(Genus), Family, Genus)
         ) |> 
  group_by(Phylum, Class, Order, Family, Genus, Compartment) |>
  summarise(mean_diff=mean(diff.btw),
            mean_disp=mean(diff.win),
            Total= n()) |> 
  group_by(Genus) |> 
   filter(n() < 2) |># View()
  ggplot() +
  geom_col(aes(x = mean_diff, 
               y = Genus ,
               fill = Compartment
  )) +
   scale_fill_manual(values = c("#00BA38","#619CFF")) +
  # facet_nested(
  #   rows = vars(Order, Family),
  #   scales = "free",
  #   space = 'free',
  #   switch = "y",
  #   nest_line = element_line(color = "black",
  #                            linetype = 2)
  # ) +
  labs(x = "Differential abundance",
       y = "Class & Order") +
  theme_light() +
  theme(legend.position = "bottom",
        strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0,
                                         colour = "black"),
        ggh4x.facet.nestline = element_line(colour = "blue",),
        strip.background = element_blank()
  )

#ggsave("figures/diff_abund_asv.svg", height = 15)
```

```{r}
aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Order, diff.btw, diff.win)) |> 
  mutate(Compartment = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column")) |> 
  group_by(Phylum, Class, Order, Compartment) |>
  summarise(mean_diff=mean(diff.btw),
            mean_disp=mean(diff.win),
            Total= n()) |> 
  filter(mean_diff > 2 | mean_diff < -2) |> 
  mutate(comb_taxa = paste(Class,":",Order, sep = "")) |> 
  group_by(Order) |>
   filter(n() < 2) |> #View()
  ggplot() +
  geom_col(aes(x = mean_diff, 
               y = comb_taxa ,
               fill = Phylum
  )) +
   #scale_fill_manual(values = c("#00BA38","#619CFF")) +
  # facet_nested(
  #   rows = vars(Order, Family),
  #   scales = "free",
  #   space = 'free',
  #   switch = "y",
  #   nest_line = element_line(color = "black",
  #                            linetype = 2)
  # ) +
  labs(x = "Differential abundance",
       y = "Class & Order") +
  theme_light() +
  theme(legend.position = "bottom",
        strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0,
                                         colour = "black"),
        ggh4x.facet.nestline = element_line(colour = "blue",),
        strip.background = element_blank()
  )

#ggsave("figures/diff_abund_asv.svg", height = 15)
```

```{r}
set.seed(3)
cols = turbo(20)[sample(1:20,20)]
# 
f3 <- aldex_full |> 
  filter(diff.btw > (diff.btw/diff.win) | diff.btw < (diff.btw/diff.win)) |> 
  rownames_to_column("asv") |> 
  left_join(asv_taxa, c("asv" = "id")) |> 
  dplyr::select(c(Phylum, Class, Order, diff.btw, diff.win)) |> 
  mutate(Compartment = ifelse(diff.btw > (diff.btw/diff.win),"Leaf","Water column")) |> 
  group_by(Phylum, Class, Order, Compartment) |>
  summarise(mean_diff=mean(diff.btw),
            mean_disp=mean(diff.win),
            Total= n()) |> 
  filter(mean_diff > 2 | mean_diff < -2) |> 
  mutate(comb_taxa = paste(Class,":",Order, sep = "")) |> 
  group_by(Order) |>
   filter(n() < 2) |> 
  ggplot() +
  geom_col(aes(x = mean_diff, 
               y = reorder(comb_taxa, mean_diff),
               fill = Phylum
  )) +
  scale_fill_viridis(discrete = TRUE, option = "turbo") +
  labs(x = "Differential abundance",
       y = "Class & Order") +
  theme_light()


#ggsave("figures/fig3.svg", height = 9.5)
```

