---
title: "CN analysis"
author: JB
date: September 2022
project: EMBRACE
output: html_notebook
---

Session Info
R version 4.2.0 (2022-04-22)
Platform: x86_64-apple-darwin17.0 (64-bit)
Running under: macOS Monterey 12.6
```{r}
sessionInfo()
packageVersion("tidyverse") # ‘1.3.1’
packageVersion("here")      # ‘1.0.1’
```

# Install and load packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("here")) install.packages("here")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("ggpubr")) install.packages("ggpubr")
```

# Create "output" folder
```{r}
here()
dir.create(here("output"))
```

# Import data
```{r}
CN_ratio <- read.table(here("Google Drive", "Marine Symbiomes Group folder","EMBRACE" ,"MS Posidonia N cycling","Biogeochemistry","Data","Raw data","CN_ratio.txt" ), header=T)
```


```{r}
CN_ratio <- read.table("./Raw data/CN_ratio.txt", header=T)
```

# Check data
```{r}
view(CN_ratio) 
names(CN_ratio)
str(CN_ratio)
```

# Reorganize table
```{r}
CN_ratio %>% 
  group_by(sample,site, incubation, treat, h) %>%
  summarise(n = n()) %>%  view()

str(CN_ratio)

CN_ratio <- CN_ratio %>% 
  mutate(
    sample = factor (x = sample, levels = c("Leaf", "Epi")),
    site = factor (x = site, levels = c("Ambient", "Vent")),
    incubation = factor (x = incubation, levels = c("light", "dark", "start")),
    treat = factor(x = treat, levels = c("Y", "N", "Ctrl"))
  ) %>% 
  view()
```

# Statistical Analysis

# C content
```{r}
# only T1= 5h

CN_ratio_C <- CN_ratio %>% 
  filter(h==5, treat!="Ctrl") %>% 
  select(-N, -CN) %>% 
  view()
```

# Leaves
```{r}
CN_C_leaf <- CN_ratio_C %>% 
  filter(sample=="Leaf", treat=="N") %>% 
  view()
```

# Outlier, normality & variance check 
```{r}
ggplot(data = CN_C_leaf, aes(x = site, y = C, color = incubation))+
  geom_boxplot(fill=NA, coef = 1.5, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# possible outlier site=ambient, incubation = dark

# remove outlier

CN_C_leaf <- CN_C_leaf %>% 
  filter(C > 33) %>% 
  view()

# test normal distribution 
qqnorm(CN_C_leaf$C)
qqline(CN_C_leaf$C)
hist(CN_C_leaf$C)

# Shapiro test for normal distribution
shapiro.test(CN_C_leaf$C)
# p-value = 0.3549 --> normally distributed 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_C_leaf$C ~ site*incubation, data = CN_C_leaf)
#p-value = 0.6212 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_C_leaf<-lm(CN_C_leaf$C~site*incubation, data=CN_C_leaf)
Anova(M_C_leaf)
# interaction not significant --> reduce model
M_C_leaf<-lm(CN_C_leaf$C~site+incubation, data=CN_C_leaf)
Anova(M_C_leaf)
summary(M_C_leaf)

# Anova Table (Type II tests)

# Response: CN_C_leaf$C
#             Sum Sq Df F value Pr(>F)
# site        0.3275  1  0.1943 0.6710
# incubation  3.9098  1  2.3205 0.1662
# Residuals  13.4791  8  

#plot model
par(mfrow=c(2,2))
plot(M_C_leaf)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_C_leaf,~site+incubation))%>%
  view()

sum <- CN_C_leaf%>% 
  group_by(incubation, site)%>% 
  summarise(sd=sd(C),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site", "incubation", "emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            incubation=incubation,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p1<-ggplot(final,aes(site,emmean,fill=incubation)) +
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  scale_fill_manual("legend", values = c("lightblue", "darkblue"), labels=c("light", "dark"))+
  theme_classic()+
  theme(legend.title = element_blank(), text = element_text(size = 20))+
  labs(title="Leaves", x="pH", y=expression(C~("%"~DW)))+
  scale_y_continuous(limits=c(0,40), breaks=c(0,5,10,15,20,25,30,35,40))
p1

ggsave(p1, filename = here("output", "Nfix_C_leaf.png"), dpi = 210, units = "cm", width = 15, height = 12)
```

# Epiphytes
```{r}
CN_C_epi <- CN_ratio_C %>% 
  filter(sample=="Epi") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_C_epi, aes(x = site, y = C, color = incubation))+
  geom_boxplot(fill=NA, coef = 1.5, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outlier

# test normal distribution 
qqnorm(CN_C_epi$C)
qqline(CN_C_epi$C)
hist(CN_C_epi$C)

# Shapiro test for normal distribution
shapiro.test(CN_C_epi$C)
# p-value = 0.04432 --> not normally distributed, still okay for Anova 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_C_epi$C ~ site*incubation, data = CN_C_epi)
#p-value = 0.002771 --> variances are different for different sites and treatments
```

# Anova
```{r}
M_C_epi<-lm(CN_C_epi$C~site*incubation, data=CN_C_epi)
# interaction significant, but factor not --> reduce model
M_C_epi<-lm(CN_C_epi$C~site+incubation, data=CN_C_epi)
Anova(M_C_epi)
summary(M_C_epi)

# Anova Table (Type II tests)

# Response: CN_C_epi$C
#             Sum Sq Df F value  Pr(>F)  
# site       116.372  1 24.8828 0.0003155 ***
# incubation  10.821  1  2.3138 0.1541357    
# Residuals   56.121 12    

#plot model
par(mfrow=c(2,2))
plot(M_C_epi)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_C_epi,~site+incubation))%>%
  view()

sum <- CN_C_epi%>% 
  group_by(incubation,site)%>% 
  summarise(sd=sd(C),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site","incubation" ,"emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            emmean=emmean,
            incubation=incubation,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p2<-ggplot(final,aes(site,emmean,fill=incubation)) +
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  scale_fill_manual("legend", values = c("lightblue", "darkblue"), labels=c("light", "dark"))+
  theme_classic()+
  theme(legend.title = element_blank(),text = element_text(size = 20))+
  labs(title="Epiphytes", x="pH", y=expression(C~("%"~DW)))+
  scale_y_continuous(limits=c(0,40), breaks=c(0,5,10,15,20,25,30,35,40))
p2

ggsave(p2, filename = here("output", "Nfix_C_epi.png"), dpi = 210, units = "cm", width = 15, height = 12)
```

# N content
```{r}
# only T1= 5h

CN_ratio_N <- CN_ratio %>% 
  filter(h==5, treat!="Ctrl") %>% 
  select(-C, -CN) %>% 
  view()
```

# Leaves
```{r}
CN_N_leaf <- CN_ratio_N %>% 
  filter(sample=="Leaf", treat=="N") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_N_leaf, aes(x = site, y = N, color = incubation))+
  geom_boxplot(fill=NA, coef = 1.5, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outliers

# test normal distribution 
qqnorm(CN_N_leaf$N)
qqline(CN_N_leaf$N)
hist(CN_N_leaf$N)

# Shapiro test for normal distribution
shapiro.test(CN_N_leaf$N)
# p-value = 0.5467 --> normally distributed 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_N_leaf$N ~ site*incubation, data = CN_N_leaf)
#p-value = 0.5487 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_N_leaf<-lm(CN_N_leaf$N~site*incubation, data=CN_N_leaf)
Anova(M_N_leaf)
# interaction not significant --> reduce model
M_N_leaf<-lm(CN_N_leaf$N~site+incubation, data=CN_N_leaf)
Anova(M_N_leaf)
summary(M_N_leaf)

# Anova Table (Type II tests)

# Response: CN_N_leaf$N
#             Sum Sq Df F value  Pr(>F)  
# site       0.02377  1  0.4044 0.5406
# incubation 0.00136  1  0.0231 0.8826
# Residuals  0.52899  9    

#plot model
par(mfrow=c(2,2))
plot(M_N_leaf)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_N_leaf,~site+incubation))%>%
  view()

sum <- CN_N_leaf%>% 
  group_by(incubation, site)%>% 
  summarise(sd=sd(N),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site", "incubation", "emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            incubation=incubation,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p3<-ggplot(final,aes(site,emmean,fill=incubation)) +
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  scale_fill_manual("legend", values = c("lightblue", "darkblue"), labels=c("light", "dark"))+
  theme_classic()+
  theme(legend.title = element_blank(),text = element_text(size = 20))+
  labs(title="Leaves", x="pH", y=expression(N~("%"~DW)))+
  scale_y_continuous(limits=c(0,2), breaks=c(0,0.25,0.5,0.75,1,1.25,1.5, 1.75, 2))
p3

ggsave(p3, filename = here("output", "Nfix_N_leaf.png"), dpi = 210, units = "cm", width = 15, height = 12)
```

# Epiphytes
```{r}
CN_N_epi <- CN_ratio_N %>% 
  filter(sample=="Epi") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_N_epi, aes(x = site, y = N, color = incubation))+
  geom_boxplot(fill=NA, coef = 2, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outlier

# test normal distribution 
qqnorm(CN_N_epi$N)
qqline(CN_N_epi$N)
hist(CN_N_epi$N)

# Shapiro test for normal distribution
shapiro.test(CN_N_epi$N)
# p-value = 0.4113 --> normally distributed 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_N_epi$N ~ site*incubation, data = CN_N_epi)
#p-value = 0.9444 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_N_epi<-lm(CN_N_epi$N~site*incubation, data=CN_N_epi)
Anova(M_N_epi)
# interaction significant, but factor not --> reduce model
M_N_epi<-lm(CN_N_epi$N~site+incubation, data=CN_N_epi)
Anova(M_N_epi)
summary(M_N_epi)

# Anova Table (Type II tests)

# Response: CN_N_epi$N
#             Sum Sq Df F value  Pr(>F)  
# site       1.90724  1 25.8068 0.0002706 ***
# incubation 0.23517  1  3.1821 0.0997370 .  
# Residuals  0.88686 12  

#plot model
par(mfrow=c(2,2))
plot(M_N_epi)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_N_epi,~site+incubation))%>%
  view()

sum <- CN_N_epi%>% 
  group_by(incubation,site)%>% 
  summarise(sd=sd(N),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site","incubation","emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            incubation=incubation,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p4<-ggplot(final,aes(site,emmean,fill=incubation)) +
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  scale_fill_manual("legend", values = c("lightblue", "darkblue"), labels=c("light", "dark"))+
  theme_classic()+
  theme(legend.title = element_blank(),text = element_text(size = 20))+
  labs(title="Epiphytes", x="pH", y=expression(N~("%"~DW)))+
  scale_y_continuous(limits=c(0,2), breaks=c(0,0.25,0.5,0.75,1,1.25,1.5, 1.75, 2))
p4

ggsave(p4, filename = here("output", "Nfix_N_epi.png"), dpi = 210, units = "cm", width = 15, height = 12)
```


# Analysis dark and light combined
since there NEVER was a difference between light and dark incubations, I combined the samples and treated them as replicates

# C content
```{r}
# only T1= 5h

CN_ratio_C <- CN_ratio %>% 
  filter(h==5, treat!="Ctrl") %>% 
  select(-N, -CN) %>% 
  view()
```

# Leaves
```{r}
CN_C_leaf <- CN_ratio_C %>% 
  filter(sample=="Leaf", treat=="N") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_C_leaf, aes(x = site, y = C, color = site))+
  geom_boxplot(fill=NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# possible outlier site=ambient 

# remove outlier

CN_C_leaf <- CN_C_leaf %>% 
  filter(C > 33) %>% 
  view()

# test normal distribution 
qqnorm(CN_C_leaf$C)
qqline(CN_C_leaf$C)
hist(CN_C_leaf$C)

# Shapiro test for normal distribution
shapiro.test(CN_C_leaf$C)
# p-value = 0.3549 --> normally distributed 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_C_leaf$C ~ site, data = CN_C_leaf)
#p-value = 0.7165 --> variances are not different for different sites 
```

# Anova
```{r}
M_C_leaf<-lm(CN_C_leaf$C~site, data=CN_C_leaf)
Anova(M_C_leaf)
summary(M_C_leaf)

# Anova Table (Type II tests)

# Response: CN_C_leaf$C
#            Sum Sq Df F value Pr(>F)
# site       0.5988  1  0.3099 0.5913
# Residuals 17.3888  9  

#plot model
par(mfrow=c(2,2))
plot(M_C_leaf)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_C_leaf,~site))%>%
  view()

sum <- CN_C_leaf%>% 
  group_by(site)%>% 
  summarise(sd=sd(C),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site", "emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p5<-ggplot(final,aes(site,emmean)) +
  geom_bar(position="dodge", stat="identity", width=0.7)+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.15)+
  theme_classic()+
  theme(text = element_text(size = 20))+
  labs(title="Leaves", x="pH", y=expression(C~("%"~DW)))+
  scale_y_continuous(limits=c(0,40), breaks=c(0,5,10,15,20,25,30,35,40))
p5

ggsave(p5, filename = here("output", "Nfix_C_leaf_combined.png"), dpi = 210, units = "cm", width = 10, height = 12)
```

# Epiphytes
```{r}
CN_C_epi <- CN_ratio_C %>% 
  filter(sample=="Epi") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_C_epi, aes(x = site, y = C, color = site))+
  geom_boxplot(fill=NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outlier

# test normal distribution 
qqnorm(CN_C_epi$C)
qqline(CN_C_epi$C)
hist(CN_C_epi$C)

# Shapiro test for normal distribution
shapiro.test(CN_C_epi$C)
# p-value = 0.04432 --> not normally distributed, still okay for Anova 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_C_epi$C ~ site, data = CN_C_epi)
#p-value = 0.08947 --> variances are not different for different sites 
```

# Anova
```{r}
M_C_epi<-lm(CN_C_epi$C~site, data=CN_C_epi)
Anova(M_C_epi)
summary(M_C_epi)

# Anova Table (Type II tests)

# Response: CN_C_epi$C
#             Sum Sq Df F value  Pr(>F)  
# site       111.928  1  21.736 0.0004446 ***
# Residuals  66.943 13    

#plot model
par(mfrow=c(2,2))
plot(M_C_epi)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_C_epi,~site))%>%
  view()

sum <- CN_C_epi%>% 
  group_by(site)%>% 
  summarise(sd=sd(C),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site","emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p6<-ggplot(final,aes(site,emmean)) +
  geom_bar(position="dodge", stat="identity", width=0.7)+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  theme_classic()+
  theme(text = element_text(size = 20))+
  labs(title="Epiphytes", x="pH", y=expression(C~("%"~DW)))+
  scale_y_continuous(limits=c(0,40), breaks=c(0,5,10,15,20,25,30,35,40))
p6

ggsave(p6, filename = here("output", "Nfix_C_epi_combined.png"), dpi = 210, units = "cm", width = 10, height = 12)
```

# N content
```{r}
# only T1= 5h

CN_ratio_N <- CN_ratio %>% 
  filter(h==5, treat!="Ctrl") %>%
  select(-C, -CN) %>% 
  view()
```

# Leaves
```{r}
CN_N_leaf <- CN_ratio_N %>% 
  filter(sample=="Leaf", treat=="N") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_N_leaf, aes(x = site, y = N, color = site))+
  geom_boxplot(fill=NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outliers

# test normal distribution 
qqnorm(CN_N_leaf$N)
qqline(CN_N_leaf$N)
hist(CN_N_leaf$N)

# Shapiro test for normal distribution
shapiro.test(CN_N_leaf$N)
# p-value = 0.5467 --> normally distributed 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_N_leaf$N ~ site, data = CN_N_leaf)
#p-value = 0.1038 --> variances are not different for different sites
```

# Anova
```{r}
M_N_leaf<-lm(CN_N_leaf$N~site, data=CN_N_leaf)
Anova(M_N_leaf)
summary(M_N_leaf)

# Anova Table (Type II tests)

# Response: CN_N_leaf$N
#            Sum Sq Df F value  Pr(>F)  
# site      0.02377  1  0.4482 0.5183
# Residuals 0.53035 10    

#plot model
par(mfrow=c(2,2))
plot(M_N_leaf)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_N_leaf,~site))%>%
  view()

sum <- CN_N_leaf%>% 
  group_by(site)%>% 
  summarise(sd=sd(N),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site", "emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p7<-ggplot(final,aes(site,emmean)) +
  geom_bar(position="dodge", stat="identity", width = 0.7)+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  theme_classic()+
  theme(text = element_text(size = 20))+
  labs(title="Leaves", x="pH", y=expression(N~("%"~DW)))+
  scale_y_continuous(limits=c(0,2), breaks=c(0,0.25,0.5,0.75,1,1.25,1.5, 1.75, 2))
p7

ggsave(p7, filename = here("output", "Nfix_N_leaf_combined.png"), dpi = 210, units = "cm", width = 10, height = 12)
```

# Epiphytes
```{r}
CN_N_epi <- CN_ratio_N %>% 
  filter(sample=="Epi") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_N_epi, aes(x = site, y = N, color = site))+
  geom_boxplot(fill=NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outlier

# test normal distribution 
qqnorm(CN_N_epi$N)
qqline(CN_N_epi$N)
hist(CN_N_epi$N)

# Shapiro test for normal distribution
shapiro.test(CN_N_epi$N)
# p-value = 0.4113 --> normally distributed 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_N_epi$N ~ site, data = CN_N_epi)
#p-value = 0.4214 --> variances are not different for different sites
```

# Anova
```{r}
M_N_epi<-lm(CN_N_epi$N~site, data=CN_N_epi)
Anova(M_N_epi)
summary(M_N_epi)

# Anova Table (Type II tests)

# Response: CN_N_epi$N
#           Sum Sq Df F value  Pr(>F)  
# site      1.8221  1  21.111 0.0005026 ***
# Residuals 1.1220 13   

#plot model
par(mfrow=c(2,2))
plot(M_N_epi)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_N_epi,~site))%>%
  view()

sum <- CN_N_epi%>% 
  group_by(site)%>% 
  summarise(sd=sd(N),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site","emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p8<-ggplot(final,aes(site,emmean)) +
  geom_bar(position="dodge", stat="identity", width = 0.7)+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  theme_classic()+
  theme(text = element_text(size = 20))+
  labs(title="Epiphytes", x="pH", y=expression(N~("%"~DW)))+
  scale_y_continuous(limits=c(0,2), breaks=c(0,0.25,0.5,0.75,1,1.25,1.5, 1.75, 2))
p8

ggsave(p8, filename = here("output", "Nfix_N_epi_combined.png"), dpi = 210, units = "cm", width = 10, height = 12)
```

# CN ratio
```{r}
# only T1= 5h

CN_ratio_CN <- CN_ratio %>% 
  filter(h==5, treat!="Ctrl") %>% 
  select(-C, -N) %>% 
  view()
```

# Leaves
```{r}
CN_CN_leaf <- CN_ratio_CN %>% 
  filter(sample=="Leaf", treat=="N") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_CN_leaf, aes(x = site, y = CN, color = site))+
  geom_boxplot(fill=NA, coef = 2, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# possible outlier site = vent

# remove outlier

CN_CN_leaf <- CN_CN_leaf %>% 
  filter(CN < 60) %>% 
  view()

# test normal distribution 
qqnorm(CN_CN_leaf$CN)
qqline(CN_CN_leaf$CN)
hist(CN_CN_leaf$CN)

# Shapiro test for normal distribution
shapiro.test(CN_CN_leaf$CN)
# p-value = 0.1247 --> normally distributed 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_CN_leaf$CN ~ site, data = CN_CN_leaf)
#p-value = 0.6676 --> variances are not different for different sites 
```

# Anova
```{r}
M_CN_leaf<-lm(CN_CN_leaf$CN~site, data=CN_CN_leaf)
Anova(M_CN_leaf)
summary(M_CN_leaf)

# Anova Table (Type II tests)

# Response: CN_CN_leaf$CN
#           Sum Sq Df F value  Pr(>F)  
# site       18.03  1  0.4565 0.5162
# Residuals 355.39  9                       

#plot model
par(mfrow=c(2,2))
plot(M_CN_leaf)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_CN_leaf,~site))%>%
  view()

sum <- CN_CN_leaf%>% 
  group_by(site)%>% 
  summarise(sd=sd(CN),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site","emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p9<-ggplot(final,aes(site,emmean)) +
  geom_bar(position="dodge", stat="identity", width = 0.7, fill="#555555")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  theme_classic()+
  theme(text = element_text(size = 15))+
  labs(title="Leaves", x="pH", y=expression(CN~ratio~(molar)))+
  scale_y_continuous(limits=c(0,45), breaks=c(0,5,10,15,20,25,30,35,40,45))
p9

ggsave(p9, filename = here("output", "Nfix_CN_ratio_leaf_combined.png"), dpi = 210, units = "cm", width = 10, height = 12)
```

# Epiphytes
```{r}
CN_CN_epi <- CN_ratio_CN %>% 
  filter(sample=="Epi") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = CN_CN_epi, aes(x = site, y = CN, color = site))+
  geom_boxplot(fill=NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(CN_CN_epi$CN)
qqline(CN_CN_epi$CN)
hist(CN_CN_epi$CN)

# Shapiro test for normal distribution
shapiro.test(CN_CN_epi$CN)
# p-value = 0.2026 --> normally distributed 

# Levene's Test for Homogeneity of Variances
leveneTest(CN_CN_epi$CN ~ site, data = CN_CN_epi)
#p-value = 0.1434 --> variances are not different for different sites 
```

# Anova
```{r}
M_CN_epi<-lm(CN_CN_epi$CN~site, data=CN_CN_epi)
Anova(M_CN_epi)
summary(M_CN_epi)

# Anova Table (Type II tests)

# Response: CN_CN_epi$CN
#           Sum Sq Df F value  Pr(>F)  
# site      198.23  1  10.526 0.006398 **
# Residuals 244.83 13   

#plot model
par(mfrow=c(2,2))
plot(M_CN_epi)
# normality and homogeneity are met
```


```{r}
# plot results from model ------------------------------------------
emm <- data.frame(emmeans(M_CN_epi,~site))%>%
  view()

sum <- CN_CN_epi%>% 
  group_by(site)%>% 
  summarise(sd=sd(CN),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site","emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

p10<-ggplot(final,aes(site,emmean)) +
  geom_bar(position="dodge", stat="identity", width = 0.7, fill="#555555")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  theme_classic()+
  theme(text = element_text(size = 15))+
  labs(title="Epiphytes", x="pH", y=element_blank())+
  scale_y_continuous(limits=c(0,45), breaks=c(0,5,10,15,20,25,30,35,40,45))
p10

ggsave(p10, filename = here("output", "Nfix_CN_ratio_epi_combined.png"), dpi = 210, units = "cm", width = 10, height = 12)
```

# create panel
```{r}
library(ggpubr)
panel1<-ggarrange(p9, p10, 
                labels = c("a", "b"),
                ncol = 2, nrow = 1, align = "hv",
                font.label = list(size = 16),
                common.legend=T,
                legend="right")
panel1

ggsave(panel1, filename = here("output", "CN_panel.png"), dpi = 210, units = "cm", width = 15, height = 8)
```
