---
title: "Inorganic Nutrient analysis"
author: JB
date: September 2022
project: EMBRACE
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Session Info
R version 4.2.0 (2022-04-22)
Platform: x86_64-apple-darwin17.0 (64-bit)
Running under: macOS Monterey 12.6
```{r}
sessionInfo()
packageVersion("tidyverse") # ‘1.3.1’
packageVersion("here")      # ‘1.0.1’
```

# Install and load packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("here")) install.packages("here")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("ggpubr")) install.packages("ggpubr")
```

# Import data
```{r}
IN <- read.table("IN.txt", header=T)
```

# Check data
```{r}
view(IN) 
names(IN)
str(IN)
```

# Reorganize table
```{r}

IN <- IN %>% 
  mutate(
    Site = factor (x = Site, levels = c("Ambient", "Vent")),
    treat1 = factor(x = treat1, levels = c("No_Epiphytes", "with_Epiphytes")),
    treat2 = factor(x = treat2, levels = c("Light", "Dark"))
  )

str(IN)
```


# Statistical Analysis

# NH4
# light incubations
```{r}
IN_light <- IN %>% 
  filter(treat2== "Light") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = IN_light, aes(x = Site, y = NH4_rate, color = treat1))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outliers

# test normal distribution 
qqnorm(IN_light$NH4_rate)
qqline(IN_light$NH4_rate)
hist(IN_light$NH4_rate)

# Shapiro test for normal distribution
shapiro.test(IN_light$NH4_rate)
# p-value = 0.1091 --> normally distributed

# Levene's Test for Homogeneity of Variances
leveneTest(IN_light$NH4_rate ~ Site*treat1, data = IN_light)
#p-value = 0.009798 --> variances are different for different sites and treatments, still okay for Anova
```

# Anova
```{r}
M_NH4_light<-lm(IN_light$NH4_rate~Site*treat1, data=IN_light)
Anova(M_NH4_light)
#interaction not significant --> reduce model
M_NH4_light<-lm(IN_light$NH4_rate~Site+treat1, data=IN_light)
Anova(M_NH4_light)
summary(M_NH4_light)

# Anova Table (Type II tests)

# Response: IN_light$NH4_rate
#           Sum Sq Df F value  Pr(>F)   
# Site      122811  1 24.9080 0.0004084 ***
# treat1      9786  1  1.9847 0.1865265    
# Residuals  54236 11  

#Adjusted R-squared:  0.6569 

#plot model
par(mfrow=c(2,2))
plot(M_NH4_light)
# normality and homogeneity are met

# posthoc test
emmeans(M_NH4_light, pairwise ~ Site+treat1)
```

# plot results 
```{r}
p1_raw<-ggplot(data = IN_light, aes(x = Site, y = NH4_rate, color=treat1))+
  geom_boxplot(fill = NA, coef = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_classic()+
  scale_color_manual(values = c("#BFD58E", "#0B6623"), labels=c("-Epi", "+Epi"))+ 
  theme(legend.title = element_blank(),text = element_text(size = 15))+
  labs(title="light", x= element_blank(), y=expression(flux~(nmol~NH[4]^+{}~cm^-2~h^-1)))+
  scale_y_continuous(limits=c(0,4.5))
p1_raw
```

# dark incubations
```{r}
IN_dark <- IN %>% 
  filter(treat2== "Dark") %>% 
  view()
```


# Outlier, normality & variance check
```{r}
ggplot(data = IN_dark, aes(x = Site, y = NH4_rate, color = treat1))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# outlier ambient, no epi

# remove outlier
IN_dark_NH4 <- IN_dark %>% 
  filter(NH4_rate > 0) %>% 
  view()

ggplot(data = IN_dark_NH4, aes(x = Site, y = NH4_rate, color = treat1))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(IN_dark_NH4$NH4_rate)
qqline(IN_dark_NH4$NH4_rate)
hist(IN_dark_NH4$NH4_rate)

# Shapiro test for normal distribution
shapiro.test(IN_dark_NH4$NH4_rate)
# p-value = 0.1218 --> normally distributed

# Levene's Test for Homogeneity of Variances
leveneTest(IN_dark_NH4$NH4_rate ~ Site*treat1, data = IN_dark_NH4)
#p-value = 0.6249 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_NH4_dark<-lm(IN_dark_NH4$NH4_rate~Site*treat1, data=IN_dark_NH4)
Anova(M_NH4_dark)
#interaction not significant --> reduce model
M_NH4_dark<-lm(IN_dark_NH4$NH4_rate~Site+treat1, data=IN_dark_NH4)
Anova(M_NH4_dark)
summary(M_NH4_dark)

# Response: IN_dark_NH4$NH4_rate
#           Sum Sq Df F value    Pr(>F)    
# Site       55769  1 60.2636 1.531e-05 ***
# treat1        55  1  0.0592    0.8126    
# Residuals   9254 10                       

#Adjusted R-squared:  0.8316  

#plot model
par(mfrow=c(2,2))
plot(M_NH4_dark)
# normality and homogeneity are met

# posthoc test
emmeans(M_NH4_dark, pairwise ~ Site+treat1)
```

# plot results 
```{r}
p2_raw<-ggplot(data = IN_dark_NH4, aes(x = Site, y = NH4_rate, color=treat1))+
  geom_boxplot(fill = NA, coef = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_classic()+
  scale_color_manual(values = c("#BFD58E", "#0B6623"), labels=c("-Epi", "+Epi"))+ 
  theme(legend.title = element_blank(),text = element_text(size = 15))+
  labs(title="dark", x=element_blank(), y=element_blank())+
  scale_y_continuous(limits=c(0, 4.5))
p2_raw
```


# NO3
# light incubations
# Outlier, normality & variance check
```{r}
ggplot(data = IN_light, aes(x = Site, y = NO3_rate, color = treat1))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# outliers ambient site

# remove outlier
IN_light_NO3 <- IN_light %>% 
  filter(NO3_rate > -0.5) %>% 
  view()

ggplot(data = IN_light_NO3, aes(x = Site, y = NO3_rate, color = treat1))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(IN_light_NO3$NO3_rate)
qqline(IN_light_NO3$NO3_rate)
hist(IN_light_NO3$NO3_rate)

# Shapiro test for normal distribution
shapiro.test(IN_light_NO3$NO3_rate)
# p-value = 0.2825 --> normally distributed

# Levene's Test for Homogeneity of Variances
leveneTest(IN_light_NO3$NO3_rate ~ Site*treat1, data = IN_light_NO3)
#p-value = 0.4448 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_NO3_light<-lm(IN_light_NO3$NO3_rate~Site*treat1, data=IN_light_NO3)
Anova(M_NO3_light)
#interaction is not significant --> reduce model
M_NO3_light<-lm(IN_light_NO3$NO3_rate~Site+treat1, data=IN_light_NO3)
Anova(M_NO3_light)
summary(M_NO3_light)

#Response: IN_light_NO3$NO3_rate
#            Sum Sq Df F value  Pr(>F)  
# Site       939.73  1  2.7747 0.1301
# treat1     650.22  1  1.9199 0.1992
# Residuals 3048.09  9 

# Adjusted R-squared:  0.1892

#plot model
par(mfrow=c(2,2))
plot(M_NO3_light)
# normality and homogeneity are met (mind the small sample size)

# posthoc test
emmeans(M_NO3_light, pairwise ~ Site+treat1)
```

# plot results 
```{r}
p3_raw<-ggplot(data = IN_light_NO3, aes(x = Site, y = NO3_rate, color=treat1))+
  geom_boxplot(fill = NA, coef = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_classic()+
  scale_color_manual(values = c("#BFD58E", "#0B6623"), labels=c("-Epi", "+Epi"))+ 
  labs(x="pH", y=expression(flux~(nmol~NO[3]^-{}~cm^-2~h^-1)))+
  theme(legend.title = element_blank(),text = element_text(size = 15))+
  scale_y_continuous(limits=c(-0.15,0.75))
p3_raw
```

# dark incubations
# Outlier, normality & variance check
```{r}
ggplot(data = IN_dark, aes(x = Site, y = NO3_rate, color = treat1))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
#  outlier vent site with epis

# remove outlier
IN_dark_NO3 <- IN_dark %>% 
  filter(NO3_rate > -0.25) %>% 
  view()

ggplot(data = IN_dark_NO3, aes(x = Site, y = NO3_rate, color = treat1))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(IN_dark_NO3$NO3_rate)
qqline(IN_dark_NO3$NO3_rate)
hist(IN_dark_NO3$NO3_rate)

# Shapiro test for normal distribution
shapiro.test(IN_dark_NO3$NO3_rate)
# p-value = 0.4851 --> normally distributed

# Levene's Test for Homogeneity of Variances
leveneTest(IN_dark_NO3$NO3_rate ~ Site*treat1, data = IN_dark_NO3)
#p-value = 0.8708 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_NO3_dark<-lm(IN_dark_NO3$NO3_rate~Site*treat1, data=IN_dark_NO3)
Anova(M_NO3_dark)
#interaction significant, but main factor not --> reduce model
M_NO3_dark<-lm(IN_dark_NO3$NO3_rate~Site+treat1, data=IN_dark_NO3)
Anova(M_NO3_dark)
summary(M_NO3_dark)

#Response: IN_dark_NO3$NO3_rate
#           Sum Sq Df F value  Pr(>F)  
# Site      2245.1  1  4.8609 0.05202 .
# treat1     894.0  1  1.9357 0.19431  
# Residuals 4618.8 10

# Adjusted R-squared:  0.2679 

#plot model
par(mfrow=c(2,2))
plot(M_NO3_dark)
# normality and homogeneity are met 

# posthoc test
emmeans(M_NO3_dark, pairwise ~ Site+treat1)
```

# plot results 
```{r}
p4_raw<-ggplot(data = IN_dark_NO3, aes(x = Site, y = NO3_rate, color=treat1))+
  geom_boxplot(fill = NA, coef = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_classic()+
  scale_color_manual(values = c("#BFD58E", "#0B6623"), labels=c("-Epi", "+Epi"))+ 
  labs(x="pH", y=element_blank())+
  theme(legend.title = element_blank(),text = element_text(size = 15))+
  scale_y_continuous(limits=c(-0.15,0.75))
p4_raw
```

# create panel
```{r}
panel1<-ggarrange(p1_raw,p2_raw,p3_raw,p4_raw, 
                labels = c("a", "b", "c", "d"),
                ncol = 2, nrow = 2, align = "hv",
                font.label = list(size = 14),
                common.legend=T,
                legend="right")
panel1

ggsave(panel1, filename = ("Nutrient_panel_raw.png"), dpi = 300, units = "cm", width = 20, height = 18)
```
