---
title: "Inorganic Nutrient analysis"
author: JB
date: September 2022
project: EMBRACE
output: html_notebook
---

Session Info
R version 4.2.0 (2022-04-22)
Platform: x86_64-apple-darwin17.0 (64-bit)
Running under: macOS Monterey 12.6
```{r}
sessionInfo()
packageVersion("tidyverse") # ‘1.3.1’
packageVersion("here")      # ‘1.0.1’
```

# Install and load packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("here")) install.packages("here")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("ggpubr")) install.packages("ggpubr")
```

# Create "output" folder
```{r}
here()
dir.create(here("output"))
```

# Import data
```{r}
IN <- read.table(here("Google Drive", "Marine Symbiomes Group folder","EMBRACE","MS Posidonia N cycling", "Biogeochemistry","Data","Raw data", "Inorganic_nutrients.txt"), header=T)
```

# Check data
```{r}
view(IN) 
names(IN)
str(IN)
```

# Reorganize table
```{r}
IN %>% 
  group_by(Experiment, Site, Incubation, Timepoint, Treatment) %>%
  summarise(n = n()) %>%  view()

IN<-IN %>% 
  filter(Experiment=="Fixation") %>% 
  rename(exp = Experiment, 
         site = Site, 
         inc = Incubation, 
         tmpt=Timepoint, 
         time = Time, treat=Treatment, code=Code) %>%
  view()

str(IN)

IN <- IN %>% 
  mutate(
    site = factor (x = site, levels = c("Ambient", "Vent"), labels = c("ambient pH", "vent pH")),
    inc = factor (x = inc, levels = c("After", "Before", "Start", "Light", "Dark", "Light/Dark")),
    treat = factor(x = treat, levels = c("Y", "N", "Ctrl")),
    tmpt = factor(x = time, levels = c("T0", "T1", "T2", "T3") )
  )
```


# Statistical Analysis

# NH4
# light incubations
# plot data
```{r}
ggplot(data = IN, aes(x = inc, y = NH4, color = site, shape = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outliers here, data looks okay (given the small sample size)
```

# calculate rates
```{r}
IN_light <- IN %>% 
  filter(inc== "Light") %>% 
  view()

# Vent pH site

IN_light_vent <- IN_light %>% 
  filter(site == "vent pH") %>% 
  view()

IN_T0_vent <- IN %>% 
  filter(inc== "Start", site =="vent pH") %>%
  summarize(NH4_mean = mean(NH4))
IN_T0_vent
# IN_T0_vent NH4 mean: 6.472667

# leaf area: 3 cm^-2

NH4_rate_light_vent<-(((((IN_light_vent$NH4)-6.472667))/IN_light_vent$time)/3)
NH4_rate_light_vent

IN_light_vent$NH4_rate<-NH4_rate_light_vent
view(IN_light_vent)

# Ambient pH site

IN_light_ambient <- IN_light %>% 
  filter(site == "ambient pH") %>%
  view()

IN_T0_ambient <- IN %>% 
  filter(inc== "Start", site =="ambient pH") %>%
  summarize(NH4_mean = mean(NH4))
IN_T0_ambient
# IN_T0_ambient NH4 mean: 5.413767

# leaf area: 3 cm^-2

NH4_rate_light_ambient<-(((((IN_light_ambient$NH4)-5.413767))/IN_light_ambient$time)/3)
NH4_rate_light_ambient

IN_light_ambient$NH4_rate<-NH4_rate_light_ambient
view(IN_light_ambient)

IN_light <- bind_rows(IN_light_vent, IN_light_ambient)%>% 
  view()


```

# Outlier, normality & variance check
```{r}
ggplot(data = IN_light, aes(x = site, y = NH4_rate, color = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# possible outlier ambient pH, treatment= Y

# remove Ctrl samples for statistical analysis

IN_light <- IN_light %>% 
  filter(treat=="Y"|treat=="N") %>% 
  view()

ggplot(data = IN_light, aes(x = site, y = NH4_rate, color = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# remove outlier

IN_light <- IN_light %>% 
  filter(NH4_rate < -0.15) %>% 
  view()


# test normal distribution 
qqnorm(IN_light$NH4_rate)
qqline(IN_light$NH4_rate)
hist(IN_light$NH4_rate)

# Shapiro test for normal distribution
shapiro.test(IN_light$NH4_rate)
# p-value = 0.2558 --> normally distributed

# Levene's Test for Homogeneity of Variances
leveneTest(IN_light$NH4_rate ~ site*treat, data = IN_light)
#p-value = 0.6578 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_NH4_light<-lm(IN_light$NH4_rate~site*treat, data=IN_light)
Anova(M_NH4_light)
#interaction not significant --> reduce model
M_NH4_light<-lm(IN_light$NH4_rate~site+treat, data=IN_light)
Anova(M_NH4_light)
summary(M_NH4_light)

# Anova Table (Type II tests)

# Response: IN_light$NH4_rate
#               Sum Sq Df F value  Pr(>F)   
# site      0.0195267  1 18.2237 0.00164 **
# treat     0.0012567  1  1.1729 0.30423   
# Residuals 0.0107150 10 


#plot model
par(mfrow=c(2,2))
plot(M_NH4_light)
# normality and homogeneity are met

# posthoc test
emmeans(M_NH4_light, pairwise ~ site+treat)
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_NH4_light,~site+treat))%>%
  view()

sum <- IN_light%>% 
  group_by(treat, site)%>% 
  summarise(sd=sd(NH4_rate),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()


final <- cbind(emm[c("site", "treat", "emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            treat=treat,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

phlabs <- c("Ambient", "Vent")

p1<-ggplot(final,aes(site,emmean,fill=treat)) +
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(0.9),width=0.2)+
  scale_fill_manual("legend", values = c("#0B6623", "#BFD58E"), labels=c("+Epi", "-Epi"))+
  theme_classic()+
  theme(legend.title = element_blank(),text = element_text(size = 15))+
  labs(title="light", x= element_blank(), y=expression(flux~(mu*mol~NH[4]^+{}~cm^-2~h^-1)))+
  scale_x_discrete(labels= phlabs)+
  scale_y_continuous(limits=c(-0.4,0), breaks=c(-0.4,-0.35,-0.3,-0.25,-0.2,-0.15,-0.1,-0.05,0))
p1

ggsave(p1, filename = here("output", "Nfix_NH4_light.png"), dpi = 210, units = "cm", width = 15, height = 12)
```

# dark incubations
# calculate rates
```{r}
IN_dark <- IN %>% 
  filter(inc== "Dark") %>% 
  view()

# Vent pH site

IN_dark_vent <- IN_dark %>% 
  filter(site == "vent pH") %>% 
  view()

IN_T0_vent <- IN %>% 
  filter(inc== "Start", site =="vent pH") %>%
  summarize(NH4_mean = mean(NH4))
IN_T0_vent
# IN_T0_vent NH4 mean: 6.472667

# leaf area: 3 cm^-2

NH4_rate_dark_vent<-(((((IN_dark_vent$NH4)-6.472667))/IN_dark_vent$time)/3)
NH4_rate_dark_vent

IN_dark_vent$NH4_rate<-NH4_rate_dark_vent
view(IN_dark_vent)


# Ambient pH site

IN_dark_ambient <- IN_dark %>% 
  filter(site == "ambient pH") %>%  
  view()

IN_T0_ambient <- IN %>% 
  filter(inc== "Start", site =="ambient pH") %>%
  summarize(NH4_mean = mean(NH4))
IN_T0_ambient
# IN_T0_ambient NH4 mean: 5.413767

# leaf area: 3 cm^-2

NH4_rate_dark_ambient<-(((((IN_dark_ambient$NH4)-5.413767))/IN_dark_ambient$time)/3)
NH4_rate_dark_ambient

IN_dark_ambient$NH4_rate<-NH4_rate_dark_ambient
view(IN_dark_ambient)

IN_dark <- bind_rows(IN_dark_vent, IN_dark_ambient)%>% 
  view()

```


# Outlier, normality & variance check
```{r}
ggplot(data = IN_dark, aes(x = site, y = NH4_rate, color = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# possible outlier ambient pH, treatment = N


# remove Ctrl samples for statistical analysis

IN_dark <- IN_dark %>% 
  filter(treat=="Y"|treat=="N") %>% 
  view()

ggplot(data = IN_dark, aes(x = site, y = NH4_rate, color = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# remove outlier
IN_dark <- IN_dark %>% 
  filter(NH4_rate < -0.1) %>% 
  view()

# test normal distribution 
qqnorm(IN_dark$NH4_rate)
qqline(IN_dark$NH4_rate)
hist(IN_dark$NH4_rate)

# Shapiro test for normal distribution
shapiro.test(IN_dark$NH4_rate)
# p-value = 0.3965 --> normally distributed

# Levene's Test for Homogeneity of Variances
leveneTest(IN_dark$NH4_rate ~ site*treat, data = IN_dark)
#p-value = 0.6249 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_NH4_dark<-lm(IN_dark$NH4_rate~site*treat, data=IN_dark)
Anova(M_NH4_dark)
#interaction not significant --> reduce model
M_NH4_dark<-lm(IN_dark$NH4_rate~site+treat, data=IN_dark)
Anova(M_NH4_dark)
summary(M_NH4_dark)

# Response: IN_dark$NH4_rate
#               Sum Sq Df F value    Pr(>F)    
# site      0.0112667  1 25.2454 0.0005183 ***
# treat     0.0000264  1  0.0592 0.8126323    
# Residuals 0.0044629 10                      


#plot model
par(mfrow=c(2,2))
plot(M_NH4_dark)
# normality and homogeneity are met

# posthoc test
emmeans(M_NH4_dark, pairwise ~ site+treat)
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_NH4_dark,~site+treat))%>%
  view()

sum <- IN_dark%>% 
  group_by(treat, site)%>% 
  summarise(sd=sd(NH4_rate),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()


final <- cbind(emm[c("site", "treat", "emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            treat=treat,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

phlabs <- c("Ambient", "Vent")

p2<-ggplot(final,aes(site,emmean,fill=treat)) +
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(.9),width=0.2)+
  scale_fill_manual("legend", values = c("#0B6623", "#BFD58E"), labels=c("+Epi", "-Epi"))+
  theme_classic()+
  theme(legend.title = element_blank(),text = element_text(size = 15))+
  labs(title="dark", x=element_blank(), y=element_blank())+
  scale_x_discrete(labels= phlabs)+
  scale_y_continuous(limits=c(-0.4,0), breaks=c(-0.4,-0.35,-0.3,-0.25,-0.2,-0.15,-0.1,-0.05,0))
p2

ggsave(p2, filename = here("output", "Nfix_NH4_dark.png"), dpi = 210, units = "cm", width = 15, height = 12)
```


# NO3
# light incubations
# plot data
```{r}
ggplot(data = IN, aes(x = inc, y = NO3, color = site, shape = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outliers here, data looks okay, but broadly ranged (mind the small sample size)
```

# calculate rates
```{r}
IN_light <- IN %>% 
  filter(inc== "Light") %>% 
  view()

# Vent pH site

IN_light_vent <- IN_light %>% 
  filter(site == "vent pH") %>% 
  view()

IN_T0_vent <- IN %>% 
  filter(inc== "Start", site =="vent pH") %>%
  summarize(NO3 = mean(NO3))
IN_T0_vent
# IN_T0_vent NO3 mean: 1.024

# leaf area: 3 cm^-2

NO3_rate_light_vent<-(((((IN_light_vent$NO3)-1.024))/IN_light_vent$time)/3)
NO3_rate_light_vent

IN_light_vent$NO3_rate<-NO3_rate_light_vent
view(IN_light_vent)

# Ambient pH site

IN_light_ambient <- IN_light %>% 
  filter(site == "ambient pH") %>%
  view()

IN_T0_ambient <- IN %>% 
  filter(inc== "Start", site =="ambient pH") %>%
  summarize(NO3_mean = mean(NO3))
IN_T0_ambient
# IN_T0_ambient NO3 mean: 0.761

# leaf area: 3 cm^-2

NO3_rate_light_ambient<-(((((IN_light_ambient$NO3)-0.761))/IN_light_ambient$time)/3)
NO3_rate_light_ambient

IN_light_ambient$NO3_rate<-NO3_rate_light_ambient
view(IN_light_ambient)

IN_light <- bind_rows(IN_light_vent, IN_light_ambient)%>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = IN_light, aes(x = site, y = NO3_rate, color = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# possible outlier ambient pH, treatment = Y + N


# remove Ctrl samples for statistical analysis

IN_light <- IN_light %>% 
  filter(treat=="Y"|treat=="N") %>% 
  view()

ggplot(data = IN_light, aes(x = site, y = NO3_rate, color = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# remove outlier
IN_light <- IN_light %>% 
  filter(NO3_rate < 0.03) %>% 
  view()

# test normal distribution 
qqnorm(IN_light$NO3_rate)
qqline(IN_light$NO3_rate)
hist(IN_light$NO3_rate)

# Shapiro test for normal distribution
shapiro.test(IN_light$NO3_rate)
# p-value = 0.275 --> normally distributed

# Levene's Test for Homogeneity of Variances
leveneTest(IN_light$NO3_rate ~ site*treat, data = IN_light)
#p-value = 0.4743 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_NO3_light<-lm(IN_light$NO3_rate~site*treat, data=IN_light)
Anova(M_NO3_light)
#interaction is not significant --> reduce model
M_NO3_light<-lm(IN_light$NO3_rate~site+treat, data=IN_light)
Anova(M_NO3_light)
summary(M_NO3_light)

# Response: IN_light$NO3_rate
#               Sum Sq Df F value Pr(>F)
# site      0.00045331  1  2.7757 0.1301
# treat     0.00031358  1  1.9201 0.1992
# Residuals 0.00146983  9

#plot model
par(mfrow=c(2,2))
plot(M_NO3_light)
# normality and homogeneity are met (mind the small sample size)

# posthoc test
emmeans(M_NO3_light, pairwise ~ site+treat)
```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_NO3_light,~site+treat))%>%
  view()

sum <- IN_light%>% 
  group_by(treat, site)%>% 
  summarise(sd=sd(NO3_rate),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site", "treat", "emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            treat=treat,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

phlabs <- c("Ambient", "Vent")

p3<-ggplot(final,aes(site,emmean,fill=treat)) +
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(.9),width=0.2)+
  scale_fill_manual("legend", values = c("#0B6623", "#BFD58E"), labels=c("+Epi", "-Epi"))+
  theme_classic()+
  theme(legend.title = element_blank(),text = element_text(size = 15))+
  labs(x="pH", y=expression(flux~(mu*mol~NO[3]^-{}~cm^-2~h^-1)))+
  scale_x_discrete(labels= phlabs)+
  scale_y_continuous(limits=c(-0.05,0.02), breaks=c(-0.05,-0.04,-0.03,-0.02,-0.01,0,0.01,0.02))
p3

ggsave(p3, filename = here("output", "Nfix_NO3_light.png"), dpi = 210, units = "cm", width = 15, height = 12)
```

# dark incubations
# calculate rates
```{r}
IN_dark <- IN %>% 
  filter(inc== "Dark") %>% 
  view()

# Vent pH site

IN_dark_vent <- IN_dark %>% 
  filter(site == "vent pH") %>% 
  view()

IN_T0_vent <- IN %>% 
  filter(inc== "Start", site =="vent pH") %>%
  summarize(NO3_mean = mean(NO3))
IN_T0_vent
# IN_T0_vent NO3 mean: 1.024

# leaf area: 3 cm^-2

NO3_rate_dark_vent<-(((((IN_dark_vent$NO3)-1.024))/IN_dark_vent$time)/3)
NO3_rate_dark_vent

IN_dark_vent$NO3_rate<-NO3_rate_dark_vent
view(IN_dark_vent)

# Ambient pH site

IN_dark_ambient <- IN_dark %>% 
  filter(site == "ambient pH") %>%  
  view()

IN_T0_ambient <- IN %>% 
  filter(inc== "Start", site =="ambient pH") %>%
  summarize(NO3_mean = mean(NO3))
IN_T0_ambient
# IN_T0_ambient NO3 mean: 0.761

# leaf area: 3 cm^-2

NO3_rate_dark_ambient<-(((((IN_dark_ambient$NO3)-0.761))/IN_dark_ambient$time)/3)
NO3_rate_dark_ambient

IN_dark_ambient$NO3_rate<-NO3_rate_dark_ambient
view(IN_dark_ambient)

IN_dark <- bind_rows(IN_dark_vent, IN_dark_ambient)%>% 
  view()

```

# Outlier, normality & variance check
```{r}
ggplot(data = IN_dark, aes(x = site, y = NO3_rate, color = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# remove Ctrl samples for statistical analysis

IN_dark <- IN_dark %>% 
  filter(treat=="Y"|treat=="N") %>% 
  view()

ggplot(data = IN_dark, aes(x = site, y = NO3_rate, color = treat))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()


# test normal distribution 
qqnorm(IN_dark$NO3_rate)
qqline(IN_dark$NO3_rate)
hist(IN_dark$NO3_rate)

# Shapiro test for normal distribution
shapiro.test(IN_dark$NO3_rate)
# p-value = 0.9998 --> normally distributed

# Levene's Test for Homogeneity of Variances
leveneTest(IN_dark$NO3_rate ~ site*treat, data = IN_dark)
#p-value = 0.582 --> variances are not different for different sites and treatments

```

# Anova
```{r}
M_NO3_dark<-lm(IN_dark$NO3_rate~site*treat, data=IN_dark)
Anova(M_NO3_dark)
#interaction not significant --> reduce model
M_NO3_dark<-lm(IN_dark$NO3_rate~site+treat, data=IN_dark)
Anova(M_NO3_dark)
summary(M_NO3_dark)

# Response: IN_dark$NO3_rate
#               Sum Sq Df F value Pr(>F)
# site      0.0007917  1  1.5799 0.2348
# treat     0.0003545  1  0.7075 0.4182
# Residuals 0.0055119 11  

#plot model
par(mfrow=c(2,2))
plot(M_NO3_dark)
# normality and homogeneity are met 

# posthoc test
emmeans(M_NO3_dark, pairwise ~ site+treat)

```

# plot results from model
```{r}
emm <- data.frame(emmeans(M_NO3_dark,~site+treat))%>%
  view()

sum <- IN_dark%>% 
  group_by(treat, site)%>% 
  summarise(sd=sd(NO3_rate),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("site", "treat", "emmean")], sum[c("sd", "se")])%>%
  summarise(site=site,
            treat=treat,
            emmean=emmean,
            sd=sd,
            se=se,
            clow= emmean-1.96*se,
            cup = emmean+1.96*se)%>% 
  view()

phlabs <- c("Ambient", "Vent")

p4<-ggplot(final,aes(site,emmean,fill=treat)) +
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(.9),width=0.2)+
  scale_fill_manual("legend", values = c("#0B6623", "#BFD58E"), labels=c("+Epi", "-Epi"))+
  theme_classic()+
  theme(legend.title = element_blank(),text = element_text(size = 15))+
  labs(x="pH", y=element_blank())+
  scale_x_discrete(labels= phlabs)+
  scale_y_continuous(limits=c(-0.05,0.02), breaks=c(-0.05,-0.04,-0.03,-0.02,-0.01,0,0.01,0.02))
p4

ggsave(p4, filename = here("output", "Nfix_NO3_dark.png"), dpi = 210, units = "cm", width = 15, height = 12)
```

# create panel
```{r}
library(ggpubr)

panel1<-ggarrange(p1,p2,p3,p4, 
                labels = c("a", "b", "c", "d"),
                ncol = 2, nrow = 2, align = "hv",
                font.label = list(size = 14),
                common.legend=T,
                legend="right")
panel1

ggsave(panel1, filename = here("output", "Nutrient_panel.png"), dpi = 300, units = "cm", width = 20, height = 18)
```
