---
title: "N2 fixation analysis"
author: JB
date: September 2022
project: EMBRACE
output: html_notebook
---

Session Info
R version 4.2.0 (2022-04-22)
Platform: x86_64-apple-darwin17.0 (64-bit)
Running under: macOS Monterey 12.6
```{r}
sessionInfo()
packageVersion("tidyverse") # ‘1.3.1’
packageVersion("here")      # ‘1.0.1’
```

# Install and load packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("here")) install.packages("here")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("ggpubr")) install.packages("ggpubr")
```

# Create "output" folder
```{r}
here()
dir.create(here("output"))
```

# Import data
```{r}
N_inc <- read.table(here("Google Drive", "Marine Symbiomes Group folder","EMBRACE" ,"MS Posidonia N cycling", "Biogeochemistry", "Data","Raw data", "N2fix_15N_incorporation.txt"), header=T)
```

# Check data
```{r}
view(N_inc) 
names(N_inc)
str(N_inc)
```

# Reorganize table
```{r}
N_inc %>% 
  group_by(Sample,Site, Phase, Treatment, Timepoint) %>%
  summarise(n = n()) %>%  view()

N_inc_new<-N_inc %>% 
  select(-Label,
         -atom_perc,
         -atom_perc_ex,
         -medium_atom_perc_ex,
         -N_ug,
         -g_DW_ind,
         -rate_ng_h) %>%
  view()

str(N_inc_new)

N_inc_new <- N_inc_new %>% 
  mutate(
    Sample=factor (x = Sample, levels = c("Epiphytes", "Leaf")),
    Site = factor (x = Site, levels = c("Ambient", "Vent")),
    Phase = factor (x = Phase, levels = c("Light", "Dark","Start")),  
    Treatment = factor (x = Treatment, levels = c("Start", "Y","N", "Ctrl")),
    Timepoint = factor (x = Timepoint, levels = c("T0", "T1","T2"))) %>% 
  view()

str(N_inc_new)
```


# Statistical Analysis

# d15N Leaves light with epiphytes
```{r}
N_leaf_L <- N_inc_new %>% 
  filter(Sample=="Leaf", Phase!="Dark", Treatment=="Y"|Treatment=="Start") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = N_leaf_L, aes(x = Site, y = delta15N, color = Timepoint, shape =Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# outlier T0, vent site, start

# remove outlier
N_leaf_L <- N_leaf_L %>% 
  filter(delta15N<10) %>% 
  view()

ggplot(data = N_leaf_L, aes(x = Site, y = delta15N, color = Timepoint, shape =Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(N_leaf_L$delta15N)
qqline(N_leaf_L$delta15N)
hist(N_leaf_L$delta15N)

# Shapiro test for normal distribution
shapiro.test(N_leaf_L$delta15N)
# p-value = 0.0888 --> normally distributed
```

# Ancova
```{r}
M_leaf<-lm(N_leaf_L$delta15N~Site*Incubation, data=N_leaf_L)
Anova(M_leaf)
# interaction not significant --> reduce model
M_leaf<-lm(N_leaf_L$delta15N~Site+Incubation, data=N_leaf_L)
Anova(M_leaf)
summary(M_leaf)

# Anova Table (Type II tests)

# Response: N_leaf_L$delta15N
#             Sum Sq Df F value Pr(>F)
# Site        0.0963  1  0.1473 0.7043
# Incubation  0.0078  1  0.0119 0.9140
# Residuals  16.9967 26 

#plot model
par(mfrow=c(2,2))
plot(M_leaf)
# normality and homogeneity are met

coef(lm(N_leaf_L$delta15N~Site+Incubation, data=N_leaf_L))
```

# plot results from model
```{r}
p1<-ggplot(N_leaf_L, aes(x=Incubation, y=delta15N, color=Site))+
  geom_point(aes(color=Site), size=2, shape=4)+
  geom_abline(intercept = 5.19939, slope = 0.00219, color="#56B4E9")+
  geom_abline(intercept = 5.19939+0.11536, slope = 0.00219, color="#E69F00")+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  theme_classic() +
  theme(legend.title = element_blank(), text = element_text(size = 16))+
  labs(title="light", x="Incubation time (h)", y=expression(delta^15*N))+
  scale_x_continuous(limits=c(0,20), breaks=c(0,5,10,15,20))+
  scale_y_continuous(limits=c(0,10), breaks=c(0,2,4,6,8,10))
p1

ggsave(p1, filename = here("output", "Nfix_d15N_leaf_L.png"), dpi = 300, units = "cm", width = 12, height = 10)
```

# d15N Leaves dark with epiphytes

```{r}
N_leaf_D <- N_inc_new %>% 
  filter(Sample=="Leaf", Phase!="Light", Treatment=="Y"|Treatment=="Start") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = N_leaf_D, aes(x = Site, y = delta15N, color = Timepoint, shape=Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# outlier T0, vent site, start

# remove outlier
N_leaf_D <- N_leaf_D %>% 
  filter(delta15N<10) %>% 
  view()

ggplot(data = N_leaf_D, aes(x = Site, y = delta15N, color = Timepoint, shape=Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(N_leaf_D$delta15N)
qqline(N_leaf_D$delta15N)
hist(N_leaf_D$delta15N)

# Shapiro test for normal distribution
shapiro.test(N_leaf_D$delta15N)
# p-value = 0.3349 --> normally distributed
```

# Ancova
```{r}
M_leaf<-lm(N_leaf_D$delta15N~Site*Incubation, data=N_leaf_D)
Anova(M_leaf)
# interaction not significant --> reduce model
M_leaf<-lm(N_leaf_D$delta15N~Site+Incubation, data=N_leaf_D)
Anova(M_leaf)
summary(M_leaf)

# Anova Table (Type II tests)

# Response: N_leaf_D$delta15N
#             Sum Sq Df F value Pr(>F)
# Site        0.2660  1  0.3294 0.5710
# Incubation  0.7965  1  0.9864 0.3298
# Residuals  20.9952 26 

#plot model
par(mfrow=c(2,2))
plot(M_leaf)
# normality and homogeneity are met
```

# plot results from model
```{r}
p2<-ggplot(N_leaf_D, aes(x=Incubation, y=delta15N, color=Site))+
  geom_point(aes(color=Site), size=2, shape=4)+
  geom_abline(intercept = 5.04409, slope = -0.02215, color="#56B4E9")+
  geom_abline(intercept = 5.04409+0.19173, slope = -0.02215, color="#E69F00")+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  theme_classic() +
  theme(legend.title = element_blank(), text = element_text(size = 16))+
  labs(title="dark", x="Incubation time (h)", y=element_blank())+
  scale_x_continuous(limits=c(0,20), breaks=c(0,5,10,15,20))+
  scale_y_continuous(limits=c(0,10), breaks=c(0,2,4,6,8,10))
p2

ggsave(p2, filename = here("output", "Nfix_d15N_leaf_D.png"), dpi = 300, units = "cm", width = 12, height = 10)
```

# create panel

```{r}
library(ggpubr)
panel1<-ggarrange(p1, p2, 
                labels = c("a", "b"),
                ncol = 2, nrow = 1, align = "hv",
                font.label = list(size = 16),
                common.legend=T,
                legend="right")
panel1
ggsave(panel1, filename = here("output", "Nfix_d15N_leaf_panel.png"), dpi = 300, units = "cm", width = 20, height = 8)
```


# d15 Epiphytes light

```{r}
N_epi_L <- N_inc_new %>% 
  filter(Sample=="Epiphytes", Phase!="Dark", Treatment=="Y"|Treatment=="Start") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = N_epi_L, aes(x = Site, y = delta15N, color = Timepoint, shape=Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# outlier T0, vent site, start

# remove outlier
N_epi_L <- N_epi_L %>% 
  filter(delta15N<17 | delta15N>18) %>% 
  view()

ggplot(data = N_epi_L, aes(x = Site, y = delta15N, color = Timepoint, shape=Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(N_epi_L$delta15N)
qqline(N_epi_L$delta15N)
hist(N_epi_L$delta15N)

# Shapiro test for normal distribution
shapiro.test(N_epi_L$delta15N)
# p-value = 3.087e-06 --> not normally distributed

# data transformation
library(LambertW)
N_epi_L$delta15N<-Gaussianize(N_epi_L$delta15N, type = "s")
qqnorm(N_epi_L$delta15N)
qqline(N_epi_L$delta15N)
hist(N_epi_L$delta15N)

# Shapiro test for normal distribution
shapiro.test(N_epi_L$delta15N)
# p-value = 0.1669 --> normally distributed
```

# Ancova
```{r}
M_epi<-lm(N_epi_L$delta15N~Site*Incubation, data=N_epi_L)
Anova(M_epi)
# interaction is significant, but site factor not --> reduce model
M_epi<-lm(N_epi_L$delta15N~Site+Incubation, data=N_epi_L)
Anova(M_epi)
summary(M_epi)

# Anova Table (Type II tests)

# Response: N_epi_L$delta15N
#             Sum Sq Df F value Pr(>F)
# Site         3.870  1  1.6216 0.2133    
# Incubation 115.565  1 48.4207 1.448e-07 ***
# Residuals   66.827 28  

#plot model
par(mfrow=c(2,2))
plot(M_epi)
# normality and homogeneity are met
```

# plot results from model
```{r}
p3<-ggplot(N_epi_L, aes(x=Incubation, y=delta15N, color=Site))+
  geom_point(aes(color=Site), size=2, shape=4)+
  geom_abline(intercept = 7.0055, slope = 0.2630, color="#56B4E9")+
  geom_abline(intercept = 7.0055+0.7223, slope = 0.2630, color="#E69F00")+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  theme_classic() +
  theme(legend.title = element_blank(), text = element_text(size = 16))+
  labs(title="light", x="Incubation time (h)", y=expression(delta^15*N))+
  scale_x_continuous(limits=c(0,20), breaks=c(0,5,10,15,20))+
  scale_y_continuous(limits=c(0,16), breaks=c(0,2,4,6,8,10,12,14,16))
p3

ggsave(p3, filename = here("output", "Nfix_d15N_epi_L.png"), dpi = 300, units = "cm", width = 12, height = 10)
```


# d15 Epiphytes dark

```{r}
N_epi_D <- N_inc_new %>% 
  filter(Sample=="Epiphytes", Phase!="Light", Treatment=="Y"|Treatment=="Start") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = N_epi_D, aes(x = Site, y = delta15N, color = Timepoint, shape=Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# outlier T0, vent site, start

# remove outlier
N_epi_D <- N_epi_D %>% 
  filter(delta15N<10) %>% 
  view()

ggplot(data = N_epi_D, aes(x = Site, y = delta15N, color = Timepoint, shape=Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(N_epi_D$delta15N)
qqline(N_epi_D$delta15N)
hist(N_epi_D$delta15N)

# Shapiro test for normal distribution
shapiro.test(N_epi_D$delta15N)
# p-value = 0.2467 --> normally distributed
```

# Ancova
```{r}
M_epi<-lm(N_epi_D$delta15N~Site*Incubation, data=N_epi_D)
Anova(M_epi)
# interaction not significant --> reduce model
M_epi<-lm(N_epi_D$delta15N~Site+Incubation, data=N_epi_D)
Anova(M_epi)
summary(M_epi)

# Anova Table (Type II tests)

# Response: N_epi_D$delta15N
#             Sum Sq Df F value Pr(>F)
# Site       1.8643  1  9.8025  0.004159 ** 
# Incubation 5.2564  1 27.6377 1.527e-05 ***
# Residuals  5.1351 27 

#plot model
par(mfrow=c(2,2))
plot(M_epi)
# normality and homogeneity are met
```

# plot results from model
```{r}
p4<-ggplot(N_epi_D, aes(x=Incubation, y=delta15N, color=Site))+
  geom_point(aes(color=Site), size=2, shape=4)+
  geom_abline(intercept = 7.83617, slope = 0.05617, color="#56B4E9")+
  geom_abline(intercept = 7.83617-0.51386, slope = 0.05617, color="#E69F00")+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  theme_classic() +
  theme(legend.title = element_blank(), text = element_text(size = 16))+
  labs(title="dark", x="Incubation time (h)", y=element_blank())+
  scale_x_continuous(limits=c(0,20), breaks=c(0,5,10,15,20))+
  scale_y_continuous(limits=c(0,16), breaks=c(0,2,4,6,8,10,12,14,16))
p4

ggsave(p4, filename = here("output", "Nfix_d15N_epi_D.png"), dpi = 300, units = "cm", width = 12, height = 10)
```

# create panel
```{r}
panel2<-ggarrange(p3, p4, 
                 labels = c("a", "b"),
                 ncol = 2, nrow = 1, align = "hv",
                 font.label = list(size = 16),
                 common.legend=T,
                 legend="right")
panel2
ggsave(panel2, filename = here("output", "Nfix_d15N_epi_panel.png"), dpi = 300, units = "cm", width = 20, height = 8)
```


# 15N incorporation rates
# only in +Epi treatments
```{r}
N_epi <- N_inc_new %>% 
  filter(Sample=="Epiphytes", Phase=="Dark"| Phase=="Light") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = N_epi, aes(x = Site, y = rate_nmol_h, color = Treatment, shape =Timepoint))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no outlier

# Controls always 0 --> removed for analysis
# only T2 (18h) considered for further analysis

N_epi_new <- N_epi %>% 
  filter(Sample=="Epiphytes", Treatment=="Y", Timepoint=="T2") %>% 
  view()

# test normal distribution 
qqnorm(N_epi_new$rate_nmol_h)
qqline(N_epi_new$rate_nmol_h)
hist(N_epi_new$rate_nmol_h)

# Shapiro test for normal distribution
shapiro.test(N_epi_new$rate_nmol_h)
# p-value = 0.0004446 -->  not normally distributed, still okay for Anova

# Levene's Test for Homogeneity of Variances
leveneTest(N_epi_new$rate_nmol_h~ Site*Phase, data = N_epi_new)
#p-value = 0.07298 --> variances are not different for different sites and treatments
```

# Anova
```{r}
M_epi<-lm(N_epi_new$rate_nmol_h~Site*Phase, data=N_epi_new)
Anova(M_epi)
summary(M_epi)

#Anova Table (Type II tests)

#Response: N_epi_new$rate_nmol_h
#            Sum Sq Df F value    Pr(>F)    
#Site        9070.5  1  7.2022 0.0199006 *  
#Phase      26109.4  1 20.7316 0.0006625 ***
#Site:Phase  9228.4  1  7.3276 0.0190628 *  
#Residuals  15112.9 12 

emm = emmeans(M_epi, specs = pairwise ~ Site:Phase)
emm

par(mfrow=c(2,2))
plot(M_epi)
# normality and homogeneity are met
```

# plot results from model
```{r}
emm<-data.frame(emmeans(M_epi,~Site*Phase))%>% 
  summarise(Site=Site,Phase=Phase,emmean=emmean)%>% 
  view()

sum <- N_epi_new%>% 
  group_by(Phase,Site)%>% 
  summarise(sd=sd(rate_nmol_h),
            n = n(),
            se=sd/sqrt(n))%>% 
  view()

final <- cbind(emm[c("Site","Phase","emmean")], sum[c("sd", "se")])%>%
  summarise(Site=Site,
            Phase=Phase,            
            emmean=emmean,
            sd=sd,
            se=se)%>% 
  view()

p5<-ggplot(data = final, aes(x = Site, y = emmean, fill=Phase))+
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(Site,ymin=emmean-se, ymax=emmean+se),position=position_dodge(.9),width=0.2)+
  scale_fill_manual(values = c("lightblue", "darkblue"), labels=c("light", "dark"))+
  theme_classic()+
  theme(legend.title = element_blank())+
  labs(x=element_blank(), y=expression({}^15*N[2]~incorporation~rate~(nmol~N~gDW^-1~h^-1)))+
  scale_y_continuous(limits=c(-1,170), breaks=c(0,25,50,75,100,125,150,175))
p5

ggsave(p5, filename = here( "output", "Nfix_15N_inc_rate.png"), dpi = 300, units = "cm", width = 15, height = 15)
```
