---
title: "Nitrification analysis"
author: JB
date: June 2022
project: EMBRACE
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Session info
R version 4.2.0 (2022-04-22)
Platform: x86_64-apple-darwin17.0 (64-bit)
Running under: macOS Monterey 12.4
```{r}
sessionInfo()
packageVersion("tidyverse") # ‘1.3.1’
packageVersion("here")      # ‘1.0.1’
```

# Install and load packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("here")) install.packages("here")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("ggpubr")) install.packages("ggpubr")
```

# Import data
```{r}
Nitr <- read.table("Nitrification.txt", header=T)
```

# Check data
```{r}
view(Nitr) 
names(Nitr)
str(Nitr)
```

# Reorganize table
```{r}
Nitr %>% 
  group_by(Site, Incubation, Treatment, Time) %>%
  summarise(n = n()) %>%  view()

str(Nitr)

Nitr_new <- Nitr %>% 
  mutate(Site = factor (x = Site, levels = c("ambient", "vent")),
    Incubation = factor (x = Incubation, levels = c("light", "dark")),
    Treatment = factor(x = Treatment, levels = c("start","Y", "N", "ctrl")))%>%
  view()
```


# Statistical Analysis

# d15N light incubations with epiphytes
```{r}
Nitr_l<- Nitr_new %>% 
  filter(Incubation=="light",
         Treatment!="N"&Treatment!="ctrl") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = Nitr_l, aes(x = Site, y = d15N.NO3.STD, color = Time,shape =Treatment))+
  geom_boxplot(fill = NA, coef = 1.5, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# outlier vent site, Treatment=Y

# remove outlier
Nitr_l <- Nitr_l %>% 
  filter(d15N.NO3.STD<30) %>% 
  view()

ggplot(data = Nitr_l, aes(x = Site, y = d15N.NO3.STD, color = Time, shape =Treatment))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()

# test normal distribution 
qqnorm(Nitr_l$d15N.NO3.STD)
qqline(Nitr_l$d15N.NO3.STD)
hist(Nitr_l$d15N.NO3.STD)

# Shapiro test for normal distribution
shapiro.test(Nitr_l$d15N.NO3.STD)
# p-value = 0.6959 --> normally distributed
```

# Ancova
```{r}
M<-lm(Nitr_l$d15N.NO3.STD~Time*Site, data=Nitr_l)
Anova(M)
# interaction not significant --> reduce model
M<-lm(Nitr_l$d15N.NO3.STD~Time+Site, data=Nitr_l)
Anova(M)
summary(M)

# Anova Table (Type II tests)

# Response: Nitr_l$d15N.NO3.STD
#            Sum Sq Df F value Pr(>F)
# Time        86.99  1  2.2332 0.1515
# Site        30.32  1  0.7784 0.3887
# Residuals   740.16 19  

#plot model
par(mfrow=c(2,2))
plot(M)
# normality and homogeneity are met

coef(lm(Nitr_l$d15N.NO3.STD~Time+Site, data=Nitr_l))
```

# plot results from model
```{r}
p1<-ggplot(Nitr_l, aes(x=Time, y=d15N.NO3.STD, color=Site))+
  geom_point(aes(color=Site), size=2, shape=4)+
  geom_abline(intercept = 0.6906 , slope = 0.3438, color="#56B4E9")+
  geom_abline(intercept = 0.6906 + 2.3587, slope = 0.3438, color="#E69F00")+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  theme_classic()+ 
  theme(legend.title = element_blank(), text = element_text(size = 16))+
  labs(title="light", x="Incubation time (h)", y=expression(delta^15*N~(NO[3]^"-")))+
  scale_y_continuous(limits=c(-10,35), breaks=c(-10,-5,0,5,10,15,20,25,30,35))+
  scale_x_continuous(limits=c(0,20), breaks=c(0,5,10,15,20))
p1

ggsave(p1, filename = here("output", "Nitr_d15N_light.png"), dpi = 300, units = "cm", width = 12, height = 10)

```


# d15N dark incubations with epiphytes
```{r}
Nitr_d<- Nitr_new %>% 
  filter(Incubation=="dark",
         Treatment!="N"&Treatment!="ctrl") %>% 
  view()
```

# Outlier, normality & variance check
```{r}
ggplot(data = Nitr_d, aes(x = Site, y = d15N.NO3.STD, color = Time,shape =Treatment))+
  geom_boxplot(fill = NA, coef = 1.5, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outliers 

# test normal distribution 
qqnorm(Nitr_d$d15N.NO3.STD)
qqline(Nitr_d$d15N.NO3.STD)
hist(Nitr_d$d15N.NO3.STD)

# Shapiro test for normal distribution
shapiro.test(Nitr_d$d15N.NO3.STD)
# p-value = 0.000996 --> not normally distributed, still okay for Anova
```

# Ancova
```{r}
M<-lm(Nitr_d$d15N.NO3.STD~Time*Site, data=Nitr_d)
Anova(M)
# interaction not significant --> reduce model
M<-lm(Nitr_d$d15N.NO3.STD~Time+Site, data=Nitr_d)
Anova(M)
summary(M)

# Anova Table (Type II tests)

# Response: Nitr_d$d15N.NO3.STD
#            Sum Sq Df F value Pr(>F)
# Time        64.09  1  0.8032 0.38645  
# Site       465.89  1  5.8381 0.03112 *
# Residuals 1037.42 13   

#plot model
par(mfrow=c(2,2))
plot(M)
# normality and homogeneity are met

coef(lm(Nitr_d$d15N.NO3.STD~Time+Site, data=Nitr_d))
```

# plot results from model
```{r}
p2<-ggplot(Nitr_d, aes(x=Time, y=d15N.NO3.STD, color=Site))+
  geom_point(aes(color=Site), size=2, shape=4)+
  geom_abline(intercept = -1.7316 , slope = 0.9506, color="#56B4E9")+
  geom_abline(intercept = -1.7316 + 11.2297, slope = 0.9506, color="#E69F00")+
  scale_color_manual(values=c("#56B4E9", "#E69F00"))+
  theme_classic()+ 
  theme(legend.title = element_blank(), text = element_text(size = 16))+
  labs(title="dark", x="Incubation time (h)", y=element_blank())+
  scale_y_continuous(limits=c(-10,35), breaks=c(-10,-5,0,5,10,15,20,25,30,35))
p2

ggsave(p2, filename = here("output", "Nitr_d15N_dark.png"), dpi = 300, units = "cm", width = 12, height = 10)
```

# create panel
```{r}
library(ggpubr)
panel<-ggarrange(p1, p2, 
                labels = c("a", "b"),
                ncol = 2, nrow = 1, align = "hv",
                font.label = list(size = 16),
                common.legend=T,
                legend="right")
panel
ggsave(panel, filename = here("output", "Nitr_d15N_panel.png"), dpi = 300, units = "cm", width = 20, height = 8)
```


# Nitrification potential
```{r}
Nitr_plot <- read.table("Nitr_graph.txt", header=T)

view(Nitr_plot) 

Nitr_plot %>% 
  group_by(Site, Incubation) %>%
  summarise(n = n()) %>%  view()

Nitr_plot_new<-Nitr_plot %>% 
  select(-Treatment) %>%
  view()

str(Nitr_plot_new)

Nitr_plot_new <- Nitr_plot_new %>% 
  mutate(
    Site = factor (x = Site, levels = c("Ambient", "Vent")),
    Incubation = factor (x = Incubation, levels = c("light", "dark"))) %>% 
  view()

str(Nitr_plot_new)
```

# Outlier, normality & variance check
```{r}
ggplot(data = Nitr_plot_new, aes(x = Site, y = rate_pmol_cm_h, color = Incubation))+
  geom_boxplot(fill = NA, coef = 3, 
               outlier.shape = "?", 
               outlier.colour = "black", 
               outlier.size = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_bw()
# no obvious outliers 

# test normal distribution 
qqnorm(Nitr_plot_new$rate_pmol_cm_h)
qqline(Nitr_plot_new$rate_pmol_cm_h)
hist(Nitr_plot_new$rate_pmol_cm_h)

# Shapiro test for normal distribution
shapiro.test(Nitr_plot_new$rate_pmol_cm_h)
# p-value = 0.002014 --> not normally distributed, still okay for Anova
```

# Anova
```{r}
M<-lm(Nitr_plot_new$rate_pmol_cm_h~Site+Incubation, data=Nitr_plot_new)
Anova(M)
summary(M)

# Anova Table (Type II tests)

# Response: Nitr_plot_new$rate_pmol_cm_h
#               Sum Sq Df F value    Pr(>F)    
# Site       0.0078867  1 66.9951 1.739e-06 ***
# Incubation 0.0007105  1  6.0358   0.02884 *  
# Residuals  0.0015304 13    

# Adjusted R-squared:  0.8256 

#plot model
par(mfrow=c(2,2))
plot(M)
# normality and homogeneity are met

emm = emmeans(M, specs = pairwise ~ Site:Incubation)
emm
```

# plot results 
```{r}
p2_raw<-ggplot(data = Nitr_plot_new, aes(x = Site, y = rate_pmol_cm_h, color=Incubation))+
  geom_boxplot(fill = NA, coef = 3)+
  geom_point(position = position_jitterdodge(), alpha = 0.5)+
  theme_classic()+
  scale_color_manual(values = c("#6F9CDE", "darkblue"), labels=c("light", "dark"))+ 
  theme(legend.title = element_blank())+
  labs(x=element_blank(), y=expression(PNR~(pmol~N~cm^-2~h^-1)))+
  scale_y_continuous(limits=c(-0.0001,0.07), breaks=c(0,0.01, 0.02,0.03,0.04,0.05,0.06,0.07))
p2_raw
```
